---
title: "Metabolomics Overfeeding Analysis"
subtitle: "Differential Abundance Analysis (DAA)"
author: "Sofie Olund Villumsen"
date: "2023"
output: 
  html_document:
    toc: true
    toc_depth: 2
---

# Initial settings
Before below analyses data has been handled (in another script):
- Imputation (KNN)
- Normalization (log2)
- Scaling (z-score/autoscale)

Clear workspace
```{r echo=FALSE}
rm(list = ls()) 
```

Load packages
```{r message=FALSE}
library(tidyverse)
library(limma) 
library(PCAtools)
library(ggrepel)
library(ggVennDiagram)
library(colorRamp2)
library(ComplexHeatmap) 
library(cowplot)
```

Set wd
```{r setup}
path <- "L:/LovbeskyttetMapper/HCOF Stem Cells/Sofie_Databehandling/github/COFpaper_omics_adiposetissue/Metabolomics"

knitr::opts_knit$set(root.dir = path)
```

Load data
```{r message=FALSE}
clinicaldata_selected <- read_csv("L:/LovbeskyttetMapper/HCOF Stem Cells/Sofie_Databehandling/Overfeeding_study/Transcriptomics/data/02_clinicaldata_selected_imputed.csv") %>% 
  mutate(id = as.numeric(id))

metabolomicdata_visitAB <- read_csv(paste(path, "data/02_visitAB_metabolomics.csv", sep="/")) 

metabolomicdata_visitAB_nafld <- read_csv(paste(path, "data/02_visitAB_nafld_metabolomics.csv", sep="/"))  

metabolite_database <- read_csv("L:/LovbeskyttetMapper/HCOF Stem Cells/Sofie_Databehandling/Baseline_study/Metabolomics/data/02_metabolite_database_info.csv") 
```


#Explorative analysis
## PCA
Create PCA object
```{r}
# Prepare data format of metabolomic data 
y_ <- metabolomicdata_visitAB %>%
  t() %>% 
  as.data.frame()
colnames(y_) <- y_[1, ]
metabolomicsdata_pca <- y_ %>% 
  filter(Sample_001_NBW_A != "Sample_001_NBW_A") %>% #stupid way of removing first row and making it the column header..
  mutate_all(., as.numeric)

# Prepare data format of clinical data 
x_ <- clinicaldata_selected %>%
  arrange(bwcat, NAFLD, visit, id) %>% 
  mutate(bwcat_visit = paste(.$bwcat, .$visit, sep= "_"),
         bwcat_visit_nafld = paste(.$bwcat, .$NAFLD, .$visit, sep= "_")) 

metadata_pca <- x_ %>% 
  select(everything(), -sample_id) %>% 
  as.data.frame()
rownames(metadata_pca) <- x_$sample_id
metadata_pca

clinicaldata_selected %>% 
  group_by(bwcat, visit) %>%
  count()

pca_object <- pca(metabolomicsdata_pca, metadata = as.data.frame(metadata_pca), removeVar = 0.1)
```


```{r}
combined_df <- x_ %>% 
  left_join(metabolomicdata_visitAB, by = join_by(sample_id)) 
```


### Biplot: bwcat, visit
```{r}
biplot_loadings_bwcat_visitAB <- biplot(pca_object, showLoadings = TRUE,
    labSize = 3, pointSize = 3, sizeLoadingsNames = 3, lab = NULL, colby = 'visit', shape = 'bwcat', 
    title = "PCA biplot", subtitle = "PC scores of samples and variable loadings", 
    # ellipse config
      ellipse = TRUE,
      ellipseType = 't',
      ellipseLevel = 0.95,
      ellipseFill = FALSE,
      ellipseAlpha = 1/4,
      ellipseLineSize = 1.0,
    legendPosition = 'right',
    xlim = c(-10,10), ylim = c(-10, 10))
biplot_loadings_bwcat_visitAB

ggsave(paste(path, "results/03_PCAbiplot_bwcat_visitAB_metabolomics.png", sep="/"), plot = biplot_loadings_bwcat_visitAB, device = "png", width = 10, height = 6)
```

### Biplot: bmi
```{r}
biplot_loadings_bmi <- biplot(pca_object, showLoadings = TRUE,
    labSize = 3, pointSize = 3, sizeLoadingsNames = 3, lab = NULL, 
    colby = 'BMI',
    title = "PCA biplot", subtitle = "PC scores of samples and variable loadings",
    legendPosition = 'right',
    xlim = c(-10,10), ylim = c(-10, 10))
biplot_loadings_bmi

ggsave(paste(path, "results/03_PCAbiplot_BMI_visitAB_metabolomics.png", sep="/"), plot = biplot_loadings_bmi, device = "png", width = 10, height = 6)
```

## Scree plot
PCs explaining the variance: Check how many PC covers 80% of the variance
```{r echo=FALSE}
which(cumsum(pca_object$variance) > 80) [1]
```

Plot
```{r warning=FALSE, echo=FALSE}
screeplot_ <- screeplot(pca_object, axisLabSize = 13, titleLabSize = 22, hline = 80, 
                        components = getComponents(pca_object, 1:20), vline = 20) +
  geom_label(aes(19, 50, label = '80% explained variation', vjust = -1, size = 8))
screeplot_

ggsave(paste(path, "results/03_screeplot_visitAB_metabolomics.png", sep="/"), plot = screeplot_, device = "png", width = 10, height = 6)
```


## Eigencorr plots 
Interpretation of r and r squared values: 
* E1: Pearson's r is used to express the correlation between two quantities. You could calculate Pearson's r to evaluate whether the two quantities are correlated. r values ranges from -1 to +1

* E2: R^2 is usually used to evaluate the quality of fit of a model on data. It expresses what fraction of the variability of your dependent variable (Y) is explained by your independent variable (X). R^2 values ranges between 0 to +1

### E1
```{r}
tiff(file = paste(path, "results/03_eigencorplot_PearsonsR_visitAB_metabolomics.tiff", sep="/"), units="in", width=8, height=13, res=300, pointsize = 0.0001)
eigencorplot(pca_object,
    components = getComponents(pca_object, 1:4),
    metavars = c("Birth weight category","Visit","Height", "Weight","BMI","Total lean mass","Total fat mass", "VAT mass","F-glucose","F-C-peptide","F-insulin", "HOMA-IR","Total cholesterol","LDL cholesterol","HDL cholesterol", "F-triglyceride","Hepatic fat","ASAT","ALAT", "GGT","NAFLD status","Birth weight - NAFLD status"),
    col = c('#283A90', 'white', '#BE2625'),
    cexCorval = 1.2,
    colCorval = 'white',
    fontCorval = 2,
    posLab = 'bottomleft',
    rotLabX = 45,
    posColKey = 'top',
    cexLabColKey = 1.5,
    scale = TRUE,
    plotRsquared = FALSE,
    main = 'PC1-6 clinical correlations Pearsons r ',
    corFUN = "pearson",
    corUSE = "pairwise.complete.obs",
    corMultipleTestCorrection = 'BH',
    signifSymbols = c('****', '***', '**', '*', ''),
    signifCutpoints = c(0, 0.0001, 0.001, 0.01, 0.05, 1),
    colFrame = 'white')
dev.off()
```

### E2
```{r}
tiff(file = paste(path, "results/03_eigencorplot_PearsonsR2_visitAB_metabolomics.tiff", sep="/"), units="in", width=8, height=13, res=300, pointsize = 0.0001)
  eigencorplot(pca_object,
    components = getComponents(pca_object, 1:4),
    metavars = c("Birth weight category","Visit","Height", "Weight","BMI","Total lean mass","Total fat mass", "VAT mass","F-glucose","F-C-peptide","F-insulin", "HOMA-IR","Total cholesterol","LDL cholesterol","HDL cholesterol", "F-triglyceride","Hepatic fat","ASAT","ALAT", "GGT","NAFLD status","Birth weight - NAFLD status"),
    col = c('white', 'gold', 'darkgreen'),
    cexCorval = 1.2,
    fontCorval = 2,
    posLab = 'bottomleft',
    posColKey = 'top',
    rotLabX = 45,
    scale = TRUE,
    main = bquote(PC1-6 ~ clinical ~ correlations ~ Pearson ~ r^2),
    plotRsquared = TRUE,
    corFUN = 'pearson',
    corUSE = 'pairwise.complete.obs',
    corMultipleTestCorrection = 'BH',
    signifSymbols = c('****', '***', '**', '*', ''),
    signifCutpoints = c(0, 0.0001, 0.001, 0.01, 0.05, 1))
dev.off()
```


## Loadingsplots
```{r echo=FALSE, warning=FALSE}
loadingsplot <- plotloadings(pca_object,
                              components = getComponents(pca_object, c(1,2,3,4)),
                              rangeRetain = 0.1,
                              labSize = 2.7,
                              absolute = FALSE,
                              title = 'Loadings plot',
                              caption = 'Top 10% variables',
                              shape = 23, shapeSizeRange = c(1, 16),
                              col = c('white', 'forestgreen'),
                              drawConnectors = FALSE)

loadingsplot
ggsave(paste(path, "results/03_PCAloadingsplot_allPC_visitAB_metabolomics.png", sep="/"), plot = loadingsplot, device = "png", width = 25, height = 11)
```

#### PC1
```{r}
loadingsplot_pc1 <- plotloadings(pca_object,
    components = getComponents(pca_object, c(1)),
    rangeRetain = 0.12, absolute = TRUE,
    col = c('black', 'green', 'forestgreen'),
    drawConnectors = TRUE, labSize = 4) + coord_flip()
loadingsplot_pc1

ggsave(paste(path, "results/03_PCAloadingsplot_PC1_visitAB_metabolomics.png", sep="/"), plot = loadingsplot_pc1, device = "png", width = 10, height = 6)
```


#### PC2
```{r}
loadingsplot_pc2 <- plotloadings(pca_object,
    components = getComponents(pca_object, c(2)),
    rangeRetain = 0.12, absolute = TRUE,
    col = c('black', 'green', 'forestgreen'),
    drawConnectors = TRUE, labSize = 4) + coord_flip()
loadingsplot_pc2

ggsave(paste(path, "results/03_PCAloadingsplot_PC2_visitAB_metabolomics.png", sep="/"), plot = loadingsplot_pc2, device = "png", width = 10, height = 6)
```

#### PC4
```{r}
loadingsplot_pc4 <- plotloadings(pca_object,
    components = getComponents(pca_object, c(4)),
    rangeRetain = 0.12, absolute = TRUE,
    col = c('black', 'green', 'forestgreen'),
    drawConnectors = TRUE, labSize = 4) + coord_flip()

loadingsplot_pc4
ggsave(paste(path, "results/03_PCAloadingsplot_PC4_visitAB_metabolomics.png", sep="/"), plot = loadingsplot_pc4, device = "png", width = 10, height = 6)
```



## Heatmap
```{r}
#OBS: Takes 3min ish to run
metabolomics_matrix <- as.matrix(metabolomicsdata_pca)
#tiff(file = paste(path, "results/03_heatmap_visitAB_metablomics.tiff", sep="/"), units="in", width=6, height=13, res=300, pointsize = 0.0001)
#heatmap(metabolomics_matrix, Colv = NA)
#dev.off()
```



# * Main analyses: Differential abundance analyses
Find which metabolites respond differently over time (between baseline and overfeeding) in LBWs relative to NBWs?

# 1. Fit model: not paired
Model for contrasts between the groups at either time point A or B
```{r}
# Count samples within the 4 groupings
metadata_pca %>% 
  group_by(bwcat, visit) %>%
  count()

# Make fit 
group <- factor(c(rep("LBW_A",26), rep("LBW_B",23), rep("NBW_A",22), rep("NBW_B",21)))
design <- model.matrix(~ 0 + group)
colnames(design) <- levels(group)
fit_ <- lmFit(metabolomicsdata_pca, design)
```

# 2. Fit model: paired
Model for contrasts between the time points A or B
```{r}
# Count samples within the 4 groupings
metadata_pca %>% 
  group_by(bwcat, visit) %>%
  count()

metadata_pca_order <- metadata_pca %>% 
  mutate(bwcat_visit = paste(bwcat, visit, sep = "_"))

# factor the group column
metadata_pca_order$bwcat_visit <- factor(metadata_pca_order$bwcat_visit)

# design
design_paired <- model.matrix(~ 0 + bwcat_visit + id, data=metadata_pca_order)
fit_paired <- lmFit(metabolomicsdata_pca, design_paired)
```



### A1: Visit (A): LBW_A vs NBW_A
```{r}
# Make contrast: The difference between the groups in response to overfeeding 
cont_A_LBWvsNBW <- makeContrasts(LBW_A-NBW_A,levels=design)
fit2_A_LBWvsNBW <- contrasts.fit(fit_, cont_A_LBWvsNBW)
fit3_A_LBWvsNBW <- eBayes(fit2_A_LBWvsNBW)

metTbl_A_LBWvsNBW <- topTable(fit3_A_LBWvsNBW, n=65, adjust.method = "fdr")
metTbl_A_LBWvsNBW
```

Direction of abundance (up or down) for significantly DE metabolites.
```{r}
metTbl_A_LBWvsNBW$topDE <- "Not significant"
metTbl_A_LBWvsNBW$topDE[metTbl_A_LBWvsNBW$logFC > 0 & metTbl_A_LBWvsNBW$P.Value < 0.05] <- "Up"
metTbl_A_LBWvsNBW$topDE[metTbl_A_LBWvsNBW$logFC < -0 & metTbl_A_LBWvsNBW$P.Value < 0.05] <- "Down"
```

#### Volcano Plot
```{r}
metTbl_A_LBWvsNBW <- metTbl_A_LBWvsNBW %>% 
  mutate(Metabolites = rownames(metTbl_A_LBWvsNBW))

ggplot(data=metTbl_A_LBWvsNBW, aes(x=logFC, y=-log10(P.Value), color=topDE, label=Metabolites)) + 
    geom_point() + 
    scale_color_manual(values=c("#283A90", "grey", "#BE2625")) + 
    labs(x = 'Log2FC', title = "Baseline (visit A): LBW vs. NBW", color = "Regulation") +
    scale_fill_discrete(breaks=c("Up", "Not significant", "Down")) +
  geom_text_repel(data = subset(metTbl_A_LBWvsNBW, topDE != "Not significant"), 
                  min.segment.length = Inf, max.overlaps = Inf, size = 3) +
    xlim(-1,1) +
  ylim(-0,3) +
  theme_classic() +
  theme(legend.position = "bottom") 
```

#### Significantly DE metabolites
FDR adjusted P-values<0.05 are considered to be DE.
```{r}
metTbl_A_LBWvsNBW_keep <- metTbl_A_LBWvsNBW$P.Value < 0.05
metTbl_A_LBWvsNBW_filtered <- metTbl_A_LBWvsNBW[metTbl_A_LBWvsNBW_keep,]
metTbl_A_LBWvsNBW_filtered
```



### A2: Visit (B): LBW_B vs NBW_B
```{r}
# Make contrast: The difference between the groups in response to overfeeding 
cont_B_LBWvsNBW <- makeContrasts(LBW_B-NBW_B,levels=design)
fit2_B_LBWvsNBW <- contrasts.fit(fit_, cont_B_LBWvsNBW)
fit3_B_LBWvsNBW <- eBayes(fit2_B_LBWvsNBW)

metTbl_B_LBWvsNBW <- topTable(fit3_B_LBWvsNBW, n=65, adjust.method = "fdr")
metTbl_B_LBWvsNBW
```

Direction of abundance (up or down) for significantly DE metabolites.
```{r}
metTbl_B_LBWvsNBW$topDE <- "Not significant"
metTbl_B_LBWvsNBW$topDE[metTbl_B_LBWvsNBW$logFC > 0 & metTbl_B_LBWvsNBW$P.Value < 0.05] <- "Up"
metTbl_B_LBWvsNBW$topDE[metTbl_B_LBWvsNBW$logFC < -0 & metTbl_B_LBWvsNBW$P.Value < 0.05] <- "Down"
```

#### Volcano Plot
```{r}
metTbl_B_LBWvsNBW <- metTbl_B_LBWvsNBW %>% 
  mutate(Metabolites = rownames(metTbl_B_LBWvsNBW))

ggplot(data = metTbl_B_LBWvsNBW, aes(x = logFC, y = -log10(P.Value), color = topDE, label = Metabolites)) +
  geom_point() +
  scale_color_manual(values = c("#283A90", "grey", "#BE2625")) +
  labs(x = 'Log2FC', title = "Overfeeding (visit B): LBW vs. NBW", color = "Regulation") +
  scale_fill_discrete(breaks = c("Up", "Not significant", "Down")) +
  geom_text_repel(data = subset(metTbl_B_LBWvsNBW, topDE != "Not significant"), 
                  min.segment.length = Inf, max.overlaps = Inf, size = 3) +
  xlim(-1, 1) +
  ylim(-0, 3) +
  theme_classic() +
  theme(legend.position = "bottom") 
```

#### Significantly DE metabolites
FDR adjusted P-values<0.05 are considered to be DE.
```{r}
metTbl_B_LBWvsNBW_keep <- metTbl_B_LBWvsNBW$P.Value < 0.05
metTbl_B_LBWvsNBW_filtered <- metTbl_B_LBWvsNBW[metTbl_B_LBWvsNBW_keep,]
metTbl_B_LBWvsNBW_filtered
```


### A3: Bwcat (LBW): LBW_B vs LBW_A 
```{r}
# Make contrast: The difference between the groups in response to overfeeding 
cont_LBW_BvsA <- makeContrasts(bwcat_visitLBW_B-bwcat_visitLBW_A,levels=design_paired)
fit2_LBW_BvsA <- contrasts.fit(fit_paired, cont_LBW_BvsA)
fit3_LBW_BvsA <- eBayes(fit2_LBW_BvsA)

metTbl_LBW_BvsA <- topTable(fit3_LBW_BvsA, n=65, adjust.method = "fdr")
metTbl_LBW_BvsA
```

```{r}
metTbl_LBW_BvsA <- topTable(fit3_LBW_BvsA, n=65, adjust.method = "fdr")
metTbl_LBW_BvsA

# For pathway analysis in IPA: Add ID's from KEGG, HMDB, PubChem and CheBI to top metabolites
metTbl_LBW_BvsA <- rownames_to_column(metTbl_LBW_BvsA, "Metabolites")
metTbl_LBW_BvsA_annotate <- left_join(x = as_tibble(metTbl_LBW_BvsA), 
                                      y = as_tibble(metabolite_database), 
                                      by = c("Metabolites" = "metabolites"))  %>%
  write_csv(paste(path, "data/03_A3_LBW_visitAB_metabolomics.csv", sep="/")) 
```

Direction of abundance (up or down) for significantly DE metabolites.
```{r}
metTbl_LBW_BvsA$topDE <- "Not significant"
metTbl_LBW_BvsA$topDE[metTbl_LBW_BvsA$logFC > 0 & metTbl_LBW_BvsA$P.Value < 0.05] <- "Up"
metTbl_LBW_BvsA$topDE[metTbl_LBW_BvsA$logFC < -0 & metTbl_LBW_BvsA$P.Value < 0.05] <- "Down"
```

#### Volcano Plot
```{r}
volcano_LBW_visitAB <- ggplot(data = metTbl_LBW_BvsA, aes(x = logFC, y = -log10(P.Value), color = topDE, label = Metabolites)) +
  geom_point() +
  scale_color_manual(values = c( "grey", "#BE2625")) +
  labs(x = 'Log2FC', title = "LBW: COF vs. baseline", color = "Regulation") +
  scale_fill_discrete(breaks = c("Up", "Not significant", "Down")) +
  geom_text_repel(data = subset(metTbl_LBW_BvsA, topDE != "Not significant"), 
                  min.segment.length = Inf, max.overlaps = Inf, size = 3) +
  theme_classic() +
  xlim(-1,1) +
  ylim(-0,3) +
  theme(legend.position = "bottom") 

volcano_LBW_visitAB

ggsave(paste(path, "results/03_volcano_A3_LBW_visitAB_metabolomics.png", sep="/"), plot = volcano_LBW_visitAB, device = "png", width = 5.5, height = 5.5)
```

#### Significantly DE metabolites
FDR adjusted P-values<0.05 are considered to be DE.
```{r}
metTbl_LBW_BvsA_keep <- metTbl_LBW_BvsA$P.Value < 0.05
metTbl_LBW_BvsA_filtered <- metTbl_LBW_BvsA[metTbl_LBW_BvsA_keep,] %>% select(Metabolites, logFC, P.Value, adj.P.Val, topDE)
metTbl_LBW_BvsA_filtered
```

### A4: Bwcat (NBW): NBW_B vs NBW_A
```{r}
# Make contrast: The difference between the groupsin response to overfeeding 
cont_NBW_BvsA <- makeContrasts(bwcat_visitNBW_B-bwcat_visitNBW_A,levels=design_paired)
fit2_NBW_BvsA <- contrasts.fit(fit_paired, cont_NBW_BvsA)
fit3_NBW_BvsA <- eBayes(fit2_NBW_BvsA)

metTbl_NBW_BvsA <- topTable(fit3_NBW_BvsA, n=65, adjust.method = "fdr")
metTbl_NBW_BvsA
```


```{r}
metTbl_NBW_BvsA <- topTable(fit3_NBW_BvsA, n=65, adjust.method = "fdr")
metTbl_NBW_BvsA

# For pathway analysis in IPA: Add ID's from KEGG, HMDB, PubChem and CheBI to top metabolites
metTbl_NBW_BvsA <- rownames_to_column(metTbl_NBW_BvsA, "Metabolites")
metTbl_NBW_BvsA_annotate <- left_join(x = as_tibble(metTbl_NBW_BvsA), 
                                      y = as_tibble(metabolite_database), 
                                      by = c("Metabolites" = "metabolites"))  %>%
  write_csv(paste(path, "data/03_A4_NBW_visitAB_metabolomics.csv", sep="/")) 
```

Direction of abundance (up or down) for significantly DE metabolites.
```{r}
metTbl_NBW_BvsA$topDE <- "Not significant"
metTbl_NBW_BvsA$topDE[metTbl_NBW_BvsA$logFC > 0 & metTbl_NBW_BvsA$P.Value < 0.05] <- "Up"
metTbl_NBW_BvsA$topDE[metTbl_NBW_BvsA$logFC < -0 & metTbl_NBW_BvsA$P.Value < 0.05] <- "Down"
```

#### Volcano Plot
```{r}
volcano_NBW_visitAB <- ggplot(data=metTbl_NBW_BvsA, aes(x=logFC, y=-log10(P.Value), color=topDE, label=Metabolites)) + 
    geom_point() + 
    scale_color_manual(values=c("#283A90", "grey", "#BE2625")) + 
    labs(x = 'Log2FC', title = "NBW: COF vs baseline", color = "Regulation") +
    scale_fill_discrete(breaks=c("Up", "Not significant", "Down")) +
  geom_text_repel(data = subset(metTbl_NBW_BvsA, topDE != "Not significant"), 
                  min.segment.length = Inf, max.overlaps = Inf, size = 3) +
    xlim(-1,1) +
  ylim(-0,3) +
  theme_classic() +
  theme() +
  theme(legend.position = "bottom") 

volcano_NBW_visitAB

ggsave(paste(path, "results/03_volcano_A4_NBW_visitAB_metabolomics.png", sep="/"), plot = volcano_NBW_visitAB, device = "png", width = 5.5, height = 5.5)
```

#### Significantly DE metabolites
FDR adjusted P-values<0.05 are considered to be DE.
```{r}
metTbl_NBW_BvsA_keep <- metTbl_NBW_BvsA$P.Value < 0.05
metTbl_NBW_BvsA_filtered <- metTbl_NBW_BvsA[metTbl_NBW_BvsA_keep,] %>% 
  select(Metabolites, logFC, P.Value, adj.P.Val, topDE)
metTbl_NBW_BvsA_filtered
```


### A5: Delta contrast LBW(AB) vs. NBW(AB)
```{r}
# Make contrast: The difference between the groups in response to overfeeding 
cont_deltaAB_LBWvsNBW <- makeContrasts((bwcat_visitLBW_B-bwcat_visitLBW_A)-(bwcat_visitNBW_B-bwcat_visitNBW_A),levels=design_paired)
fit2_deltaAB_LBWvsNBW <- contrasts.fit(fit_paired, cont_deltaAB_LBWvsNBW)
fit3_deltaAB_LBWvsNBW <- eBayes(fit2_deltaAB_LBWvsNBW)

metTbl_deltaAB_LBWvsNBW <- topTable(fit3_deltaAB_LBWvsNBW, n=65, adjust.method = "fdr")
metTbl_deltaAB_LBWvsNBW
```


```{r}
metTbl_deltaAB_LBWvsNBW <- topTable(fit3_deltaAB_LBWvsNBW, n=65, adjust.method = "fdr")
metTbl_deltaAB_LBWvsNBW

# For pathway analysis in IPA: Add ID's from KEGG, HMDB, PubChem and CheBI to top metabolites
metTbl_deltaAB_LBWvsNBW <- rownames_to_column(metTbl_deltaAB_LBWvsNBW, "Metabolites")
metTbl_deltaAB_LBWvsNBW_annotate <- left_join(x = as_tibble(metTbl_deltaAB_LBWvsNBW), 
                                      y = as_tibble(metabolite_database), 
                                      by = c("Metabolites" = "metabolites"))  %>%
  write_csv(paste(path, "data/03_A5_deltaAB_LBWvsNBW_metabolomics.csv", sep="/")) 
```

Direction of abundance (up or down) for significantly DE metabolites.
```{r}
metTbl_deltaAB_LBWvsNBW$topDE <- "Not significant"
metTbl_deltaAB_LBWvsNBW$topDE[metTbl_deltaAB_LBWvsNBW$logFC > 0 & metTbl_deltaAB_LBWvsNBW$P.Value < 0.05] <- "Up"
metTbl_deltaAB_LBWvsNBW$topDE[metTbl_deltaAB_LBWvsNBW$logFC < 0 & metTbl_deltaAB_LBWvsNBW$P.Value < 0.05] <- "Down"

# Add the significant lipids to the plot
metTbl_deltaAB_LBWvsNBW$DElabel <- NA
metTbl_deltaAB_LBWvsNBW$DElabel[metTbl_deltaAB_LBWvsNBW$topDE != "Not significant"] <- metTbl_deltaAB_LBWvsNBW$Metabolites[metTbl_deltaAB_LBWvsNBW$topDE != "Not significant"]
```

####Volcano plot
```{r}
# Define the desired order of the variables
desired_order <- c("Up", "Down", "Not significant")
volcano_delta_visitAB_NBWvsLBW <- 
  ggplot(data=metTbl_deltaAB_LBWvsNBW, aes(x=logFC, 
                                           y=-log10(P.Value), 
                                           color=topDE,
                                           label=DElabel)) + 
  geom_point() + 
  theme_classic() + 
  theme(legend.position = "bottom") +
  geom_text_repel(min.segment.length = Inf, max.overlaps = Inf, size = 4) +
  scale_color_manual(values = c("Up" = "#BE2625", "Down" = "#283A90", "Not significant" = "grey"), 
                     breaks = desired_order, 
                     labels = desired_order) +
  xlim(-1.3,1.3) +
  ylim(-0,3) +
  labs(x = 'LogFC', color = "Regulation", title = "Delta changes: LBW vs NBW") 
volcano_delta_visitAB_NBWvsLBW


ggsave(paste(path, "results/03_volcano_A5_delta_visitAB_NBWvsLBW_metabolomics.png", sep="/"), plot = volcano_delta_visitAB_NBWvsLBW, device = "png", width = 7.5, height = 5)
```


#### Significantly DE metabolites
FDR adjusted P-values<0.05 are considered to be DE.
```{r}
metTbl_deltaAB_LBWvsNBW_keep <- metTbl_deltaAB_LBWvsNBW$P.Value < 0.05
metTbl_deltaAB_LBWvsNBW_filtered <- metTbl_deltaAB_LBWvsNBW[metTbl_deltaAB_LBWvsNBW_keep,] %>% select(Metabolites, logFC, P.Value, adj.P.Val, topDE)
metTbl_deltaAB_LBWvsNBW_filtered
```

#### Boxplot w/o
```{r}
# List of bile acids
sigmet_fdr <- metTbl_deltaAB_LBWvsNBW_filtered %>%
  filter(P.Value < 0.05) %>% 
  select(Metabolites) 

sigmet_fdr <- as.list(sigmet_fdr)

# Create a list to store the plots
plot_list <- list()

# Loop over significant bile acids to create plots and store them in the list
for (i in 1:length(sigmet_fdr$Metabolites)) {
  variable <- sigmet_fdr$Metabolites[i]
  title <- sigmet_fdr$Metabolites[i]
  
  p <- combined_df %>%
    ggplot(aes(x = bwcat_visit, y = !!sym(variable), fill = bwcat)) + 
    geom_boxplot(outlier.shape = 1, alpha = 0.4) +
    scale_fill_manual(values = c("LBW" = "yellow", "NBW" = "forestgreen")) +
    geom_jitter(aes(shape = NAFLD), position = position_jitter(width = 0.2), alpha = 0.5) +
    theme_classic() +
    theme(legend.position = "right", axis.text.x = element_text(angle = 90, hjust = 1)) +
    labs(title = title) +
    ylab("Conc. (log2-scaled values)") +
    xlab("Birth Weight Groups")
  
  plot_list[[i]] <- p
}

# Combine the plots using the cowplot package
combined_plots_1 <- plot_grid(plotlist = plot_list[1:4])
combined_plots_1

combined_plots_2 <- plot_grid(plotlist = plot_list[5:8])
combined_plots_2
```



# * Subgroup Analyses: NAFLD
## PCA
Prepare data
```{r}
#Metabolomics data
# Transpose the data frame 
yy_nafld <- t(metabolomicdata_visitAB_nafld) %>%
  as.data.frame() %>% 
  mutate_all(~ str_replace(., "_0+", "_"))

# Set the column names to the first row of the transposed data
colnames(yy_nafld) <- yy_nafld[1, ] 

# Remove the first row from the data. Also, convert all values to numeric
metabolomicsdata_pca_nafld <- yy_nafld %>%
  slice(-1) %>%  # Skip the first row
  mutate_all(as.numeric)  # Convert all columns to numeric

#Clinical data
# Arrange the data frame 
x_nafld <- clinicaldata_selected %>%
  arrange(bwcat, NAFLD, visit, id) %>% 
  mutate(sample_id = paste("Sample", id, bwcat, visit, NAFLD, sep = "_"))

# Prepare metadata for PCA by excluding the 'sample_id' column
metadata_pca_nafld <- x_nafld %>% 
  select(-sample_id) %>% 
  as.data.frame()

# Set the row names of the metadata data frame to the 'sample_id' values
rownames(metadata_pca_nafld) <- x_nafld$sample_id

# Display the metadata_pca_nafld data frame
metadata_pca_nafld
```

Create PCA object
```{r}
clinicaldata_selected %>% 
  group_by(bwcat, NAFLD, visit) %>%
  count()

pca_object_nafld <- pca(metabolomicsdata_pca_nafld, metadata = as.data.frame(metadata_pca_nafld), removeVar = 0.1)
```


#### Biplot: nafld, visit
```{r}
biplot_loadings_bwcat_visitAB_nafld <- biplot(pca_object_nafld, showLoadings = TRUE,
    labSize = 3, pointSize = 3, sizeLoadingsNames = 3, lab = NULL, colby = 'NAFLD', shape = 'visit', 
    title = "PCA biplot", subtitle = "PC scores of samples and variable loadings", 
    # ellipse config
      ellipse = TRUE,
      ellipseType = 't',
      ellipseLevel = 0.95,
      ellipseFill = FALSE,
      ellipseAlpha = 1/4,
      ellipseLineSize = 1.0,
    legendPosition = 'right',
    xlim = c(-10,10), ylim = c(-10, 10))
biplot_loadings_bwcat_visitAB_nafld

ggsave(paste(path, "results/03_PCAbiplot_nafld_visitAB__metabolomics.png", sep="/"), plot = biplot_loadings_bwcat_visitAB_nafld, device = "png", width = 10, height = 6)
```

Which metabolites respond differently over time (between baseline and overfeeding) in LBWs w/ and w/o NAFLD relative to NBWs?

# 1. Fit model: not paired
```{r}
# Count samples within the 6 groupings
metadata_pca_nafld %>% 
  group_by(bwcat, NAFLD, visit) %>%
  count()

metadata_pca_nafld_order <- metadata_pca_nafld %>% 
  mutate(bwcat_nafld_visit = paste(.$bwcat, .$NAFLD, .$visit, sep = "_")) %>% 
  mutate(bwcat_nafld_visit_ = case_when(bwcat_nafld_visit == "LBW_FALSE_A" ~ "LBW_woNAFLD_A",
                                        bwcat_nafld_visit == "LBW_TRUE_A" ~ "LBW_wNAFLD_A",
                                        bwcat_nafld_visit == "LBW_FALSE_B" ~ "LBW_woNAFLD_B",
                                        bwcat_nafld_visit == "LBW_TRUE_B" ~ "LBW_wNAFLD_B",
                                        bwcat_nafld_visit == "NBW_FALSE_A" ~ "NBW_woNAFLD_A",
                                        bwcat_nafld_visit == "NBW_FALSE_B" ~ "NBW_woNAFLD_B")) %>% 
  mutate(sample_id = paste("Sample", .$id, .$bwcat_nafld_visit_, sep="_")) %>% 
  arrange(bwcat, NAFLD, visit, id) %>% 
  select(sample_id, bwcat_nafld_visit_, BMI, id)

# factor the group column
metadata_pca_nafld_order$bwcat_nafld_visit_ <- factor(metadata_pca_nafld_order$bwcat_nafld_visit_)

design_nafld <- model.matrix(~ 0 + bwcat_nafld_visit_ + BMI, data=metadata_pca_nafld_order)
colnames(design_nafld) <- c("LBWwNAFLD_A", "LBWwNAFLD_B", "LBWwoNAFLD_A", "LBWwoNAFLD_B", "NBWwoNAFLD_A", "NBWwoNAFLD_B", "BMI")
fit_nafld_ <- lmFit(metabolomicsdata_pca_nafld, design_nafld)
```

# 2. Fit model: paired
```{r}
design_nafld_paired <- model.matrix(~ 0 + bwcat_nafld_visit_ + BMI + id, data=metadata_pca_nafld_order)
fit_nafld_paired <- lmFit(metabolomicsdata_pca_nafld, design_nafld_paired)
```


## A6: Visit (A): LBWwoNAFLD_A vs NBW_A
```{r}
# Make contrast: The difference between the groupsin response to overfeeding 
cont_A_LBWwoNAFLDvsNBW <- makeContrasts(LBWwoNAFLD_A-NBWwoNAFLD_A,levels=design_nafld)
fit2_A_LBWwoNAFLDvsNBW <- contrasts.fit(fit_nafld_, cont_A_LBWwoNAFLDvsNBW)
fit3_A_LBWwoNAFLDvsNBW <- eBayes(fit2_A_LBWwoNAFLDvsNBW)

metTbl_A_LBWwoNAFLDvsNBW <- topTable(fit3_A_LBWwoNAFLDvsNBW, n=65, adjust.method = "fdr")
metTbl_A_LBWwoNAFLDvsNBW
```

```{r}
metTbl_A_LBWwoNAFLDvsNBW$topDE <- "Not significant"
metTbl_A_LBWwoNAFLDvsNBW$topDE[metTbl_A_LBWwoNAFLDvsNBW$logFC > 0 & metTbl_A_LBWwoNAFLDvsNBW$P.Value < 0.05] <- "Up"
metTbl_A_LBWwoNAFLDvsNBW$topDE[metTbl_A_LBWwoNAFLDvsNBW$logFC < -0 & metTbl_A_LBWwoNAFLDvsNBW$P.Value < 0.05] <- "Down"
```

### Volcano plot
```{r}
metTbl_A_LBWwoNAFLDvsNBW <- metTbl_A_LBWwoNAFLDvsNBW %>% 
  mutate(Metabolites = rownames(metTbl_A_LBWwoNAFLDvsNBW))

ggplot(data=metTbl_A_LBWwoNAFLDvsNBW, aes(x=logFC, y=-log10(P.Value), color=topDE, label=Metabolites)) + 
    geom_point() + 
    scale_color_manual(values=c("#283A90", "grey", "#BE2625")) + 
    labs(x = 'Log2FC', title = "Baseline (visit A): LBWwoNAFLD vs. NBW", color = "Regulation") +
    geom_text_repel(data = subset(metTbl_A_LBWwoNAFLDvsNBW, topDE != "Not significant"), 
                    min.segment.length = Inf, max.overlaps = Inf, size = 3) +
    scale_fill_discrete(breaks=c("Up", "Not significant", "Down")) +
    xlim(-1,1) +
  ylim(-0,3) +
  theme_classic() +
  theme(legend.position = "bottom")

```

### Significantly DE metabolites
FDR adjusted P-values<0.05 are considered to be DE.
```{r}
metTbl_A_LBWwoNAFLDvsNBW_keep <- metTbl_A_LBWwoNAFLDvsNBW$P.Value < 0.05
metTbl_A_LBWwoNAFLDvsNBW_filtered <- metTbl_A_LBWwoNAFLDvsNBW[metTbl_A_LBWwoNAFLDvsNBW_keep,] %>% select(logFC, P.Value,adj.P.Val, topDE)
metTbl_A_LBWwoNAFLDvsNBW_filtered
```


## A7: Visit (A): LBWwNAFLD_A vs NBW_A
```{r}
# Make contrast: The difference between the groupsin response to overfeeding 
cont_A_LBWwNAFLDvsNBW <- makeContrasts(LBWwNAFLD_A-NBWwoNAFLD_A,levels=design_nafld)
fit2_A_LBWwNAFLDvsNBW <- contrasts.fit(fit_nafld_, cont_A_LBWwNAFLDvsNBW)
fit3_A_LBWwNAFLDvsNBW <- eBayes(fit2_A_LBWwNAFLDvsNBW)

metTbl_A_LBWwNAFLDvsNBW <- topTable(fit3_A_LBWwNAFLDvsNBW, n=65, adjust.method = "fdr")
metTbl_A_LBWwNAFLDvsNBW
```

Direction of abundance (up or down) for significantly DE metabolites.
```{r}
metTbl_A_LBWwNAFLDvsNBW$topDE <- "Not significant"
metTbl_A_LBWwNAFLDvsNBW$topDE[metTbl_A_LBWwNAFLDvsNBW$logFC > 0 & metTbl_A_LBWwNAFLDvsNBW$P.Value < 0.05] <- "Up"
metTbl_A_LBWwNAFLDvsNBW$topDE[metTbl_A_LBWwNAFLDvsNBW$logFC < -0 & metTbl_A_LBWwNAFLDvsNBW$P.Value < 0.05] <- "Down"
```

#### Volcano Plot
```{r}
metTbl_A_LBWwNAFLDvsNBW <- metTbl_A_LBWwNAFLDvsNBW %>% 
  mutate(Metabolites = rownames(metTbl_A_LBWwNAFLDvsNBW))

ggplot(data=metTbl_A_LBWwNAFLDvsNBW, aes(x=logFC, y=-log10(P.Value), color=topDE,label=Metabolites)) + 
    geom_point() + 
    scale_color_manual(values=c("#283A90", "grey", "#BE2625")) + 
    labs(x = 'Log2FC', title = "Baseline (visit A): LBWwNAFLD vs. NBW", color = "Regulation") +
    scale_fill_discrete(breaks=c("Up", "Not significant", "Down")) +
  geom_text_repel(data = subset(metTbl_A_LBWwNAFLDvsNBW, topDE != "Not significant"), 
                  min.segment.length = Inf, max.overlaps = Inf, size = 3) +
    xlim(-2,2) +
  ylim(-0,3) +
  theme_classic() + 
  theme(legend.position = "bottom")
```



#### Significantly DE metabolites
FDR adjusted P-values<0.05 are considered to be DE.
```{r}
metTbl_A_LBWwNAFLDvsNBW_keep <- metTbl_A_LBWwNAFLDvsNBW$P.Value < 0.05
metTbl_A_LBWwNAFLDvsNBW_filtered <- metTbl_A_LBWwNAFLDvsNBW[metTbl_A_LBWwNAFLDvsNBW_keep,] %>% select(logFC, P.Value,adj.P.Val, topDE)
metTbl_A_LBWwNAFLDvsNBW_filtered
```


## A8: Visit (B): LBWwoNAFLD_B vs NBW_B
```{r}
# Make contrast: The difference between the groupsin response to overfeeding 
cont_B_LBWwoNAFLDvsNBW <- makeContrasts(LBWwoNAFLD_B-NBWwoNAFLD_B,levels=design_nafld)
fit2_B_LBWwoNAFLDvsNBW <- contrasts.fit(fit_nafld_, cont_B_LBWwoNAFLDvsNBW)
fit3_B_LBWwoNAFLDvsNBW <- eBayes(fit2_B_LBWwoNAFLDvsNBW)

metTbl_B_LBWwoNAFLDvsNBW <- topTable(fit3_B_LBWwoNAFLDvsNBW, n=65, adjust.method = "fdr")
metTbl_B_LBWwoNAFLDvsNBW
```

Direction of abundance (up or down) for significantly DE metabolites.
```{r}
metTbl_B_LBWwoNAFLDvsNBW$topDE <- "Not significant"
metTbl_B_LBWwoNAFLDvsNBW$topDE[metTbl_B_LBWwoNAFLDvsNBW$logFC > 0 & metTbl_B_LBWwoNAFLDvsNBW$P.Value < 0.05] <- "Up"
metTbl_B_LBWwoNAFLDvsNBW$topDE[metTbl_B_LBWwoNAFLDvsNBW$logFC < -0 & metTbl_B_LBWwoNAFLDvsNBW$P.Value < 0.05] <- "Down"
```

#### Volcano Plot
```{r}
metTbl_B_LBWwoNAFLDvsNBW <- metTbl_B_LBWwoNAFLDvsNBW %>% 
  mutate(Metabolites = rownames(metTbl_B_LBWwoNAFLDvsNBW))

ggplot(data=metTbl_B_LBWwoNAFLDvsNBW, aes(x=logFC, y=-log10(P.Value), color=topDE, label=Metabolites)) + 
    geom_point() + 
    scale_color_manual(values=c("#283A90", "grey", "#BE2625")) + 
    labs(x = 'Log2FC', title = "Overfeeding (visit B): LBWwoNAFLD vs. NBW") +
    scale_fill_discrete(breaks=c("Up", "Not significant", "Down")) +
  geom_text_repel(data = subset(metTbl_B_LBWwoNAFLDvsNBW, topDE != "Not significant"), 
                  min.segment.length = Inf, max.overlaps = Inf, size = 3) +
  xlim(-1.5,1.5) +
  ylim(-0,3) +
  theme_classic() + 
  theme(legend.position = "bottom")
```



#### Significantly DE metabolites
FDR adjusted P-values<0.05 are considered to be DE.
```{r}
metTbl_B_LBWwoNAFLDvsNBW_keep <- metTbl_B_LBWwoNAFLDvsNBW$P.Value < 0.05
metTbl_B_LBWwoNAFLDvsNBW_filtered <- metTbl_B_LBWwoNAFLDvsNBW[metTbl_B_LBWwoNAFLDvsNBW_keep,] %>% select(logFC, P.Value,adj.P.Val, topDE)
metTbl_B_LBWwoNAFLDvsNBW_filtered
```

## A9: Visit (B): LBWwNAFLD_B vs NBW_B
```{r}
# Make contrast: The difference between the groupsin response to overfeeding 
cont_B_LBWwNAFLDvsNBW <- makeContrasts(LBWwNAFLD_B-NBWwoNAFLD_B,levels=design_nafld)
fit2_B_LBWwNAFLDvsNBW <- contrasts.fit(fit_nafld_, cont_B_LBWwNAFLDvsNBW)
fit3_B_LBWwNAFLDvsNBW <- eBayes(fit2_B_LBWwNAFLDvsNBW)

metTbl_B_LBWwNAFLDvsNBW <- topTable(fit3_B_LBWwNAFLDvsNBW, n=65, adjust.method = "fdr")
metTbl_B_LBWwNAFLDvsNBW
```


Direction of abundance (up or down) for significantly DE metabolites.
```{r}
metTbl_B_LBWwNAFLDvsNBW$topDE <- "Not significant"
metTbl_B_LBWwNAFLDvsNBW$topDE[metTbl_B_LBWwNAFLDvsNBW$logFC > 0 & metTbl_B_LBWwNAFLDvsNBW$P.Value < 0.05] <- "Up"
metTbl_B_LBWwNAFLDvsNBW$topDE[metTbl_B_LBWwNAFLDvsNBW$logFC < -0 & metTbl_B_LBWwNAFLDvsNBW$P.Value < 0.05] <- "Down"
```

#### Volcano Plot
```{r}
metTbl_B_LBWwNAFLDvsNBW <- metTbl_B_LBWwNAFLDvsNBW %>% 
  mutate(Metabolites = rownames(metTbl_B_LBWwNAFLDvsNBW))

ggplot(data=metTbl_B_LBWwNAFLDvsNBW, aes(x=logFC, y=-log10(P.Value), color=topDE, label=Metabolites)) + 
    geom_point() + 
    scale_color_manual(values=c("grey", "#BE2625")) + 
    labs(x = 'Log2FC', title = "Overfeeding (visit B): LBWwNAFLD vs. NBW") +
    scale_fill_discrete(breaks=c("Up", "Not significant", "Down")) +
  geom_text_repel(data = subset(metTbl_B_LBWwNAFLDvsNBW, topDE != "Not significant"), 
                  min.segment.length = Inf, max.overlaps = Inf, size = 3) +
   xlim(-1.5,1.5) +
  ylim(-0,3) +
  theme_classic() + 
  theme(legend.position = "bottom")
```





#### Significantly DE metabolites
FDR adjusted P-values<0.05 are considered to be DE.
```{r}
metTbl_B_LBWwNAFLDvsNBW_keep <- metTbl_B_LBWwNAFLDvsNBW$P.Value < 0.05
metTbl_B_LBWwNAFLDvsNBW_filtered <- metTbl_B_LBWwNAFLDvsNBW[metTbl_B_LBWwNAFLDvsNBW_keep,] %>% select(logFC, P.Value,adj.P.Val, topDE)
metTbl_B_LBWwNAFLDvsNBW_filtered
```



## A10: Bwcat_NAFLD (LBWwoNAFLD): LBWwoNAFLD_B vs LBWwoNAFLD_A
```{r}
# Make contrast: The difference between the groupsin response to overfeeding 
cont_LBWwoNAFLD_BvsA <- makeContrasts(bwcat_nafld_visit_LBW_woNAFLD_B-bwcat_nafld_visit_LBW_woNAFLD_A,levels=design_nafld_paired)
fit2_LBWwoNAFLD_BvsA <- contrasts.fit(fit_nafld_paired, cont_LBWwoNAFLD_BvsA)
fit3_LBWwoNAFLD_BvsA <- eBayes(fit2_LBWwoNAFLD_BvsA)

metTbl_LBWwoNAFLD_BvsA <- topTable(fit3_LBWwoNAFLD_BvsA, n=65, adjust.method = "fdr")
metTbl_LBWwoNAFLD_BvsA
```

```{r}
# For pathway analysis in IPA: Add ID's from KEGG, HMDB, PubChem and CheBI to top metabolites
metTbl_LBWwoNAFLD_BvsA <- rownames_to_column(metTbl_LBWwoNAFLD_BvsA, "Metabolites")
metTbl_LBWwoNAFLD_BvsA_annotate <- left_join(x = as_tibble(metTbl_LBWwoNAFLD_BvsA), 
                                      y = as_tibble(metabolite_database), 
                                      by = c("Metabolites" = "metabolites"))  %>%
  write_csv(paste(path, "data/03_A10_LBWwoNAFLD_visitAB_metabolomics.csv", sep="/")) 
```


Direction of abundance (up or down) for significantly DE metabolites.
```{r}
metTbl_LBWwoNAFLD_BvsA$topDE <- "Not significant"
metTbl_LBWwoNAFLD_BvsA$topDE[metTbl_LBWwoNAFLD_BvsA$logFC > 0 & metTbl_LBWwoNAFLD_BvsA$P.Value < 0.05] <- "Up"
metTbl_LBWwoNAFLD_BvsA$topDE[metTbl_LBWwoNAFLD_BvsA$logFC < -0 & metTbl_LBWwoNAFLD_BvsA$P.Value < 0.05] <- "Down"
```

#### Volcano Plot
```{r}
ggplot(data=metTbl_LBWwoNAFLD_BvsA, aes(x=logFC, y=-log10(P.Value), color=topDE, label=Metabolites)) + 
    geom_point() + 
    scale_color_manual(values=c("grey", "#BE2625")) + 
    labs(x = 'Log2FC', title = "LBWwoNAFLD: COF vs. baseline") +
    scale_fill_discrete(breaks=c("Up", "Not significant", "Down")) +
  geom_text_repel(data = subset(metTbl_LBWwoNAFLD_BvsA, topDE != "Not significant"), 
                  min.segment.length = Inf, max.overlaps = Inf, size = 3) +
   xlim(-1.5,1.5) +
  ylim(-0,3.2) +
  theme_classic() + 
  theme(legend.position = "bottom")
```

#### Significantly DE metabolites
FDR adjusted P-values<0.05 are considered to be DE.
```{r}
metTbl_LBWwoNAFLD_BvsA_keep <- metTbl_LBWwoNAFLD_BvsA$P.Value < 0.05
metTbl_LBWwoNAFLD_BvsA_filtered <- metTbl_LBWwoNAFLD_BvsA[metTbl_LBWwoNAFLD_BvsA_keep,] %>% select(logFC, P.Value,adj.P.Val, topDE)
metTbl_LBWwoNAFLD_BvsA_filtered
```



## A11: Bwcat_NAFLD (LBWwNAFLD): LBWwNAFLD_B vs LBWwNAFLD_A
```{r}
# Make contrast: The difference between the groupsin response to overfeeding 
cont_LBWwNAFLD_BvsA <- makeContrasts(bwcat_nafld_visit_LBW_wNAFLD_B-bwcat_nafld_visit_LBW_wNAFLD_A,levels=design_nafld_paired)
fit2_LBWwNAFLD_BvsA <- contrasts.fit(fit_nafld_paired, cont_LBWwNAFLD_BvsA)
fit3_LBWwNAFLD_BvsA <- eBayes(fit2_LBWwNAFLD_BvsA)

metTbl_LBWwNAFLD_BvsA <- topTable(fit3_LBWwNAFLD_BvsA, n=65, adjust.method = "fdr")
metTbl_LBWwNAFLD_BvsA 
```

```{r}
# For pathway analysis in IPA: Add ID's from KEGG, HMDB, PubChem and CheBI to top metabolites
metTbl_LBWwNAFLD_BvsA <- rownames_to_column(metTbl_LBWwNAFLD_BvsA, "Metabolites")
metTbl_LBWwNAFLD_BvsA_annotate <- left_join(x = as_tibble(metTbl_LBWwNAFLD_BvsA), 
                                      y = as_tibble(metabolite_database), 
                                      by = c("Metabolites" = "metabolites"))  %>%
  write_csv(paste(path, "data/03_A11_LBWwNAFLD_visitAB_metabolomics.csv", sep="/")) 
```

Direction of abundance (up or down) for significantly DE metabolites.
```{r}
metTbl_LBWwNAFLD_BvsA$topDE <- "Not significant"
metTbl_LBWwNAFLD_BvsA$topDE[metTbl_LBWwNAFLD_BvsA$logFC > 0 & metTbl_LBWwNAFLD_BvsA$P.Value < 0.05] <- "Up"
metTbl_LBWwNAFLD_BvsA$topDE[metTbl_LBWwNAFLD_BvsA$logFC < -0 & metTbl_LBWwNAFLD_BvsA$P.Value < 0.05] <- "Down"
```

#### Volcano Plot
```{r}
ggplot(data=metTbl_LBWwNAFLD_BvsA, aes(x=logFC, y=-log10(P.Value), color=topDE, label=Metabolites)) + 
    geom_point() + 
    scale_color_manual(values=c("grey", "#BE2625")) + 
    labs(x = 'Log2FC', title = "LBWwNAFLD: COF vs. baseline") +
      scale_fill_discrete(breaks=c("Up", "Not significant", "Down")) +
  geom_text_repel(data = subset(metTbl_LBWwNAFLD_BvsA, topDE != "Not significant"), 
                  min.segment.length = Inf, max.overlaps = Inf, size = 3) +
   xlim(-2,2) +
  ylim(-0,3) +
  theme_classic() + 
  theme(legend.position = "bottom")
```

#### Significantly DE metabolites
FDR adjusted P-values<0.05 are considered to be DE.
```{r}
metTbl_LBWwNAFLD_BvsA_keep <- metTbl_LBWwNAFLD_BvsA$P.Value < 0.05
metTbl_LBWwNAFLD_BvsA_filtered <- metTbl_LBWwNAFLD_BvsA[metTbl_LBWwNAFLD_BvsA_keep,] %>% select(Metabolites, logFC, P.Value,adj.P.Val, topDE)
metTbl_LBWwNAFLD_BvsA_filtered 
```


## A12: Delta contrast LBWwoNAFLD(AB) vs. NBW(AB)
```{r}
# Make contrast: The difference between the groupsin response to overfeeding 
cont_deltaAB_LBWwoNAFLDvsNBW <- makeContrasts(
  (bwcat_nafld_visit_LBW_woNAFLD_B-bwcat_nafld_visit_LBW_woNAFLD_A)-
    (bwcat_nafld_visit_NBW_woNAFLD_B-bwcat_nafld_visit_NBW_woNAFLD_A),
                                              levels=design_nafld_paired)
fit2_deltaAB_LBWwoNAFLDvsNBW <- contrasts.fit(fit_nafld_paired, cont_deltaAB_LBWwoNAFLDvsNBW)
fit3_deltaAB_LBWwoNAFLDvsNBW <- eBayes(fit2_deltaAB_LBWwoNAFLDvsNBW)

metTbl_deltaAB_LBWwoNAFLDvsNBW <- topTable(fit3_deltaAB_LBWwoNAFLDvsNBW, n=65, adjust.method = "fdr")
metTbl_deltaAB_LBWwoNAFLDvsNBW
```


```{r}
# For pathway analysis in IPA: Add ID's from KEGG, HMDB, PubChem and CheBI to top metabolites
metTbl_deltaAB_LBWwoNAFLDvsNBW <- rownames_to_column(metTbl_deltaAB_LBWwoNAFLDvsNBW, "Metabolites")
metTbl_deltaAB_LBWwoNAFLDvsNBW_annotate <- left_join(x = as_tibble(metTbl_deltaAB_LBWwoNAFLDvsNBW), 
                                      y = as_tibble(metabolite_database), 
                                      by = c("Metabolites" = "metabolites"))  %>%
  write_csv(paste(path, "data/03_A12_deltaAB_LBWwoNAFLDvsNBW_metabolomics_imputed.csv", sep="/")) 
```

Direction of abundance (up or down) for significantly DE metabolites.
```{r}
metTbl_deltaAB_LBWwoNAFLDvsNBW$topDE <- "Not significant"
metTbl_deltaAB_LBWwoNAFLDvsNBW$topDE[metTbl_deltaAB_LBWwoNAFLDvsNBW$logFC > 0 & metTbl_deltaAB_LBWwoNAFLDvsNBW$P.Value < 0.05] <- "Up"
metTbl_deltaAB_LBWwoNAFLDvsNBW$topDE[metTbl_deltaAB_LBWwoNAFLDvsNBW$logFC < 0 & metTbl_deltaAB_LBWwoNAFLDvsNBW$P.Value < 0.05] <- "Down"

# Add the significant lipids to the plot
metTbl_deltaAB_LBWwoNAFLDvsNBW$DElabel <- NA
metTbl_deltaAB_LBWwoNAFLDvsNBW$DElabel[metTbl_deltaAB_LBWwoNAFLDvsNBW$topDE != "Not significant"] <- metTbl_deltaAB_LBWwoNAFLDvsNBW$Metabolites[metTbl_deltaAB_LBWwoNAFLDvsNBW$topDE != "Not significant"]
```

#### Volcano Plot
```{r}
volcano_delta_visitAB_NBWvsLBWwoNAFLD_metabolomics <- ggplot(data=metTbl_deltaAB_LBWwoNAFLDvsNBW, aes(x=logFC, y=-log10(P.Value), color=topDE, label=DElabel)) + 
  
  geom_point() + 
  theme_classic() + 
  theme(legend.position = "bottom") +
  geom_text_repel(min.segment.length = Inf, max.overlaps = Inf, size = 4) +
  scale_color_manual(values = c("Up" = "#BE2625", "Down" = "#283A90", "Not significant" = "grey"), 
                     breaks = desired_order, 
                     labels = desired_order) +
  xlim(-2.2,2.2) +
  ylim(-0,3) +
  labs(x = 'LogFC', color = "Regulation", title = "Delta changes: LBW w/o NAFLD vs NBW") 

volcano_delta_visitAB_NBWvsLBWwoNAFLD_metabolomics

ggsave(paste(path, "results/03_volcano_A12_delta_visitAB_NBWvsLBWwoNAFLD_metabolomics.png", sep="/"), plot = volcano_delta_visitAB_NBWvsLBWwoNAFLD_metabolomics, device = "png", width = 5.5, height = 5)
```

#### Significantly DE metabolites
FDR adjusted P-values<0.05 are considered to be DE.
```{r}
metTbl_deltaAB_LBWwoNAFLDvsNBW_keep <- metTbl_deltaAB_LBWwoNAFLDvsNBW$P.Value < 0.05
metTbl_deltaAB_LBWwoNAFLDvsNBW_filtered <- metTbl_deltaAB_LBWwoNAFLDvsNBW[metTbl_deltaAB_LBWwoNAFLDvsNBW_keep,] %>% select(Metabolites, logFC, P.Value, adj.P.Val, topDE)
metTbl_deltaAB_LBWwoNAFLDvsNBW_filtered
```



## A13: Delta contrast LBWwNAFLD(AB) vs. NBW(AB)
```{r}
# Make contrast: The difference between the groupsin response to overfeeding 
cont_deltaAB_LBWwNAFLDvsNBW <- makeContrasts(
  (bwcat_nafld_visit_LBW_wNAFLD_B-bwcat_nafld_visit_LBW_wNAFLD_A)-
    (bwcat_nafld_visit_NBW_woNAFLD_B-bwcat_nafld_visit_NBW_woNAFLD_A),levels=design_nafld_paired)
fit2_deltaAB_LBWwNAFLDvsNBW <- contrasts.fit(fit_nafld_paired, cont_deltaAB_LBWwNAFLDvsNBW)
fit3_deltaAB_LBWwNAFLDvsNBW <- eBayes(fit2_deltaAB_LBWwNAFLDvsNBW)

metTbl_deltaAB_LBWwNAFLDvsNBW <- topTable(fit3_deltaAB_LBWwNAFLDvsNBW, n=65, adjust.method = "fdr")
metTbl_deltaAB_LBWwNAFLDvsNBW
```

```{r}
# For pathway analysis in IPA: Add ID's from KEGG, HMDB, PubChem and CheBI to top metabolites
metTbl_deltaAB_LBWwNAFLDvsNBW <- rownames_to_column(metTbl_deltaAB_LBWwNAFLDvsNBW, "Metabolites")
metTbl_deltaAB_LBWwNAFLDvsNBW_annotate <- left_join(x = as_tibble(metTbl_deltaAB_LBWwNAFLDvsNBW), 
                                      y = as_tibble(metabolite_database), 
                                      by = c("Metabolites" = "metabolites"))  %>%
  write_csv(paste(path, "data/03_A13_deltaAB_LBWwNAFLDvsNBW_metabolomics_imputed.csv", sep="/")) 
```

Direction of abundance (up or down) for significantly DE metabolites.
```{r}
metTbl_deltaAB_LBWwNAFLDvsNBW$topDE <- "Not significant"
metTbl_deltaAB_LBWwNAFLDvsNBW$topDE[metTbl_deltaAB_LBWwNAFLDvsNBW$logFC > 0 & metTbl_deltaAB_LBWwNAFLDvsNBW$P.Value < 0.05] <- "Up"
metTbl_deltaAB_LBWwNAFLDvsNBW$topDE[metTbl_deltaAB_LBWwNAFLDvsNBW$logFC < 0 & metTbl_deltaAB_LBWwNAFLDvsNBW$P.Value < 0.05] <- "Down"

# Add the significant lipids to the plot
metTbl_deltaAB_LBWwNAFLDvsNBW$DElabel <- NA
metTbl_deltaAB_LBWwNAFLDvsNBW$DElabel[metTbl_deltaAB_LBWwNAFLDvsNBW$topDE != "Not significant"] <- metTbl_deltaAB_LBWwNAFLDvsNBW$Metabolites[metTbl_deltaAB_LBWwNAFLDvsNBW$topDE != "Not significant"]
```

#### Volcano Plot
```{r}
volcano_delta_visitAB_NBWvsLBWwNAFLD_metabolomics <- ggplot(data=metTbl_deltaAB_LBWwNAFLDvsNBW, aes(x=logFC, y=-log10(P.Value), color=topDE, label=DElabel)) + 
  
  geom_point() + 
  theme_classic() + 
  theme(legend.position = "bottom") +
  geom_text_repel(min.segment.length = Inf, max.overlaps = Inf, size = 4) +
  scale_color_manual(values = c("Up" = "#BE2625", "Down" = "#283A90", "Not significant" = "grey"), 
                     breaks = desired_order, 
                     labels = desired_order) +
  xlim(-2,2.2) +
  ylim(-0,3.3) +
  labs(x = 'LogFC', color = "Regulation", title = "Delta changes: LBW w/ NAFLD vs NBW") 

volcano_delta_visitAB_NBWvsLBWwNAFLD_metabolomics

ggsave(paste(path, "results/03_volcano_A13_delta_visitAB_NBWvsLBWwNAFLD_metabolomics.png", sep="/"), plot = volcano_delta_visitAB_NBWvsLBWwNAFLD_metabolomics, device = "png", width = 5.5, height = 5)
```

#### Significantly DE metabolites
FDR adjusted P-values<0.05 are considered to be DE.
```{r}
metTbl_deltaAB_LBWwNAFLDvsNBW_keep <- metTbl_deltaAB_LBWwNAFLDvsNBW$P.Value < 0.05
metTbl_deltaAB_LBWwNAFLDvsNBW_filtered <- metTbl_deltaAB_LBWwNAFLDvsNBW[metTbl_deltaAB_LBWwNAFLDvsNBW_keep,] %>% select(Metabolites, logFC, P.Value, adj.P.Val, topDE)
metTbl_deltaAB_LBWwNAFLDvsNBW_filtered
```




# Heatmaps for paper
```{r}
A5_topmet <- metTbl_deltaAB_LBWvsNBW_filtered %>% 
  mutate(comparison = "A5_deltaAB_LBWvsNBW") %>% 
  select(Metabolites, comparison, logFC, P.Value, adj.P.Val, topDE) %>% 
  rename(Regulation = topDE) %>% 
  arrange(Metabolites) %>% 
  pivot_wider(
    id_cols = Metabolites,
    names_from = comparison,
    values_from = logFC
  )

# Create a list of data frames (x) and a list of names (y)
x <- list(metTbl_LBW_BvsA, 
          metTbl_NBW_BvsA, 
          metTbl_LBWwoNAFLD_BvsA, 
          metTbl_LBWwNAFLD_BvsA)  
y <- list("A3_LBW_BvsA", "A4_NBW_BvsA", "A10_LBWwoNAFLD_BvsA", "A11_LBWwNAFLD_BvsA") 

result_list <- list()

# Loop over the data frames in x and characters in y
for (i in seq_along(x)) {
  result <- x[[i]] %>%
    mutate(comparison = y[[i]]) %>%
    select(Metabolites, comparison, logFC, P.Value, adj.P.Val, topDE) %>%
    arrange(Metabolites) %>%
    pivot_wider(
      id_cols = Metabolites,
      names_from = comparison,
      values_from = c(logFC, P.Value)
    )
  
  # Name the result data frame after y[i]
  result_name <- y[[i]]
  result_list[[result_name]] <- result
}

# Combine the individual data frames into one table
combined_df_topmet <- result_list[[1]] 
for (i in 2:length(result_list)) {
  combined_df_topmet <- left_join(combined_df_topmet, result_list[[i]], by = "Metabolites")
}

topmet_combined_df <- A5_topmet %>% 
  left_join(combined_df_topmet, by = join_by(Metabolites)) %>% 
  select(everything(), -c(Metabolites, A5_deltaAB_LBWvsNBW)) %>% 
  as.data.frame()
rownames(topmet_combined_df) <- A5_topmet$Metabolites
topmet_combined_df

topmet_combined_df %>% 
  select(contains("P.Value"))
```

```{r}
topmet_combined_df_logfc <- topmet_combined_df %>% 
  select(contains("logFC")) %>% 
  select(logFC_A4_NBW_BvsA, logFC_A3_LBW_BvsA, logFC_A10_LBWwoNAFLD_BvsA, logFC_A11_LBWwNAFLD_BvsA)


heatmap_obj <- Heatmap(topmet_combined_df_logfc, 
                       col = colorRamp2(c(min(topmet_combined_df_logfc), 0, max(topmet_combined_df_logfc)), 
                                        c("#283A90", "white", "#BE2625")),
                       name = "Log2 \nFold Changes",
                       cluster_rows = FALSE,  # To prevent row clustering
                       column_names_side = "top",
                       row_names_side = "left",
                       show_column_names = TRUE,  # To show column names
                       column_names_rot = 0,
                       column_names_centered = TRUE,
                       column_labels = c("NBW", "LBW All","LBW w/o NAFLD", "LBW w/ NAFLD"),
                       show_row_names = TRUE, # To show row names
                       cluster_columns = FALSE,
                       column_names_gp = gpar(fontsize = 8),
                       row_names_gp = gpar(fontsize = 11), 
                       show_heatmap_legend = TRUE
                       ) 

heatmap_obj

# Save the heatmap as a PDF file
tiff(file = paste(path, "results/03_heatmap_deltametabolites_A3_A4_A10_A11_metabolomics.tiff", sep="/"),
    width = 20,
    height = 10,
    units = "cm",
    res = 600)

print(heatmap_obj)
dev.off()
```
# Boxplot w/ nafld
```{r}
# List of bile acids
sigmet_fdr <- metTbl_deltaAB_LBWvsNBW_filtered %>%
  filter(P.Value < 0.05) %>% 
  select(Metabolites) 

sigmet_fdr <- as.list(sigmet_fdr)

# Create a list to store the plots
plot_list <- list()

# Loop over significant bile acids to create plots and store them in the list
for (i in 1:length(sigmet_fdr$Metabolites)) {
  variable <- sigmet_fdr$Metabolites[i]
  title <- sigmet_fdr$Metabolites[i]
  
  p <- combined_df %>%
    ggplot(aes(x = bwcat_visit_nafld, y = !!sym(variable), fill = bwcat)) + 
    geom_boxplot(outlier.shape = 1, alpha = 0.4) +
    scale_fill_manual(values = c("LBW" = "yellow", "NBW" = "forestgreen")) +
    geom_jitter(aes(shape = NAFLD), position = position_jitter(width = 0.2), alpha = 0.5) +
    theme_classic() +
    theme(legend.position = "right", axis.text.x = element_text(angle = 90, hjust = 1)) +
    labs(title = title) +
    ylab("Conc. (log2-scaled values)") +
    xlab("Birth Weight Groups")
  
  plot_list[[i]] <- p
}

# Combine the plots using the cowplot package
combined_plots_1 <- plot_grid(plotlist = plot_list[1:4])
combined_plots_1

combined_plots_2 <- plot_grid(plotlist = plot_list[5:8])
combined_plots_2
```