---
title: "Transcriptomics COF study"
subtitle: "Pathway analysis"
author: "Sofie Olund Villumsen"
date: "2023"
output: 
  html_document:
    toc: true
    toc_depth: 2
---

# Initial settings
Clear workspace
```{r echo=FALSE}
rm(list = ls()) 
```

Load packages
```{r message=FALSE, warning=FALSE}
library(ggsankey)
library(networkD3)
library(RColorBrewer)
library(tidyverse)
library(webshot)
```

Set wd
```{r setup}
path <- "L:/LovbeskyttetMapper/HCOF Stem Cells/Sofie_Databehandling/github/COFpaper_3omicslayers/Transcriptomics"

knitr::opts_knit$set(root.dir = path)
```

Load data
```{r message=FALSE}
enricher_lbw <- read_tsv(paste(path, "pathway_analysis/Enricher_output/04_A3_LBW_BvsA_SAT_RNAseq_proteincoding_Pathway_Reactome_2022_table_paired.txt", sep = "/"))

enricher_nbw <- read_tsv(paste(path, "pathway_analysis/Enricher_output/04_A4_NBW_BvsA_SAT_RNAseq_proteincoding_Pathway_Reactome_2022_table_paired.txt", sep = "/"))

enricher_delta <- read_tsv(paste(path, "pathway_analysis/Enricher_output/04_A5_deltachanges_SAT_RNAseq_proteincoding_Pathway_Reactome_2022_table.txt", sep = "/"))

enricher_lbwwoNAFLD <- read_tsv(paste(path, "pathway_analysis/Enricher_output/04_A10_LBWwoNAFLD_BvsA_SAT_RNAseq_proteincoding_Pathway_Reactome_2022_table_paired.txt", sep = "/"))

enricher_lbwwNAFLD <- read_tsv(paste(path, "pathway_analysis/Enricher_output/04_A11_LBWwNAFLD_BvsA_SAT_RNAseq_proteincoding_Pathway_Reactome_2022_table_paired.txt", sep = "/"))

enricher_deltaLBWwoNAFLDvsNBW <- read_tsv(paste(path, "pathway_analysis/Enricher_output/04_A12_deltachanges_LBWwoNAFLDvsNBW_SAT_RNAseq_proteincoding_Pathway_Reactome_2022_table.txt", sep = "/"))

enricher_deltaLBWwNAFLDvsNBW <- read_tsv(paste(path, "pathway_analysis/Enricher_output/04_A13_deltachanges_LBWwNAFLDvsNBW_SAT_RNAseq_proteincoding_Pathway_Reactome_2022_table.txt", sep = "/"))
```

# A3: LBW
## Sankey plot
```{r}
# Make Sankey df
sankeydf_enricher_lbw <- enricher_lbw %>%
  filter(`Adjusted P-value` < 0.05) %>% 
  select(Term, Genes, `Adjusted P-value`) %>% 
  separate_rows(Genes, sep = ";") %>% 
  separate(Term, into = c("Pathways"), sep = "R-HSA") 
sankeydf_enricher_lbw
```


Nodes: LBW
```{r}
# Genes unique
df_lbw_unique_genes <- sankeydf_enricher_lbw %>%
  distinct(Genes) %>%
  mutate(ID = Genes, 
         x = 1, 
         y = row_number())

# Pathways unique
df_lbw_unique_pathways <- sankeydf_enricher_lbw %>%
  distinct(Pathways) %>%
  mutate(ID = Pathways, 
         x = 2, 
         y = row_number())

# Concatenate df's
nodes_lbw <- bind_rows(df_lbw_unique_genes, df_lbw_unique_pathways) %>% 
  select(ID) %>% as.data.frame() %>% 
  mutate(ID_nr = row_number() -1) 
nodes_lbw
```

Edges: LBW
```{r}
edges_lbw <- sankeydf_enricher_lbw %>%
  mutate(gene_type = Genes) %>% 
  left_join(nodes_lbw, by = c("Genes" = "ID")) %>% 
  left_join(nodes_lbw, by = c("Pathways" = "ID")) %>% 
  mutate(source = ID_nr.x,
         target = ID_nr.y,
         value = -log10(`Adjusted P-value`)) %>% 
  select(source, target, value, gene_type)

edges_lbw
```

Create sankey plot: LBW
```{r}
# Create the network graph
network_lbw <- sankeyNetwork(
  Links = edges_lbw,
  Nodes = nodes_lbw,
  Source = "source",
  Target = "target",
  Value = "value",
  NodeID = "ID",
  LinkGroup = "gene_type", # hash out to color nodes
  fontSize = 70,
  nodeWidth = 40,
  NodeGroup = NULL, # hash out to color nodes
  sinksRight = FALSE,
  colourScale = JS("d3.scaleOrdinal(d3.schemeCategory20);"),
  iterations = 0
)

network_lbw

# Save file paths
savesankey_html_lbw <- paste(path, "pathway_analysis/sankeyplot/04_sankeyplot_topgenes_pathways_A3_LBW_BvsA_proteincoding_SAT_rnaseq.html", sep = "/")
savesankey_jpeg_lbw <- paste(path, "pathway_analysis/sankeyplot/04_sankeyplot_topgenes_pathways_A3_LBW_BvsA_proteincoding_SAT_rnaseq.jpeg", sep = "/")

# Save network as html
saveNetwork(network_lbw, savesankey_html_lbw)

# Save network as png
webshot(savesankey_html_lbw, savesankey_jpeg_lbw, zoom = 3,
        vwidth = 3000,
        vheight = 2000)
```



# A4: NBW
## Sankey plot
```{r}
# Make Sankey df
sankeydf_enricher_nbw <- enricher_nbw %>%
  filter(`Adjusted P-value` < 0.05) %>% 
  select(Term, Genes,`Adjusted P-value`) %>% 
  separate_rows(Genes, sep = ";") %>% 
  separate(Term, into = c("Pathways"), sep = "R-HSA") 
sankeydf_enricher_nbw
```

Nodes: NBW
```{r}
# Genes unique
df_nbw_unique_genes <- sankeydf_enricher_nbw %>%
  distinct(Genes) %>%
  mutate(ID = Genes, 
         x = 1, 
         y = row_number())

# Pathways unique
df_nbw_unique_pathways <- sankeydf_enricher_nbw %>%
  distinct(Pathways) %>%
  mutate(ID = Pathways, 
         x = 2, 
         y = row_number())

# Concatenate df's
nodes_nbw <- bind_rows(df_nbw_unique_genes, df_nbw_unique_pathways) %>% 
  select(ID) %>% as.data.frame() %>% 
  mutate(ID_nr = row_number() -1) 
nodes_nbw
```


Edges: NBW
```{r}
edges_nbw <- sankeydf_enricher_nbw %>%
  mutate(gene_type = Genes) %>% 
  left_join(nodes_nbw, by = c("Genes" = "ID")) %>% 
  left_join(nodes_nbw, by = c("Pathways" = "ID")) %>% 
  mutate(source = ID_nr.x,
         target = ID_nr.y,
         value = -log10(`Adjusted P-value`)) %>% 
  select(source, target, value, gene_type)

edges_nbw
```


Create the sankey plot: NBW
```{r}
# Create the network graph
network_nbw <- sankeyNetwork(
  Links = edges_nbw,
  Nodes = nodes_nbw,
  Source = "source",
  Target = "target",
  Value = "value",
  NodeID = "ID",
  fontSize = 70,
  nodeWidth = 40,
  LinkGroup = "gene_type", # hash out to color nodes
  NodeGroup = NULL, # hash out to color nodes
  sinksRight = FALSE,
  colourScale = JS("d3.scaleOrdinal(d3.schemeCategory20);"),
  iterations = 0
)

network_nbw

# Save file paths
savesankey_html_nbw <- paste(path, "pathway_analysis/sankeyplot/04_sankeyplot_topgenes_pathways_A4_NBW_BvsA_proteincoding_SAT_rnaseq.html", sep = "/")
savesankey_jpeg_nbw <- paste(path, "pathway_analysis/sankeyplot/04_sankeyplot_topgenes_pathways_A4_NBW_BvsA_proteincoding_SAT_rnaseq.jpeg", sep = "/")

# Save network as html
saveNetwork(network_nbw, savesankey_html_nbw)

# Save network as png
webshot(savesankey_html_nbw, savesankey_jpeg_nbw, zoom = 3,
        vwidth = 3000,
        vheight = 2000)
```


# A5: Delta
## Sankey plot
```{r}
# Make Sankey df
sankeydf_enricher_delta <- enricher_delta %>%
  filter(`Adjusted P-value` < 0.05) %>% 
  select(Term, Genes,`Adjusted P-value`) %>% 
  separate_rows(Genes, sep = ";") %>% 
  separate(Term, into = c("Pathways"), sep = "R-HSA") 
sankeydf_enricher_delta
```


Nodes: LBW
```{r}
# Genes unique
df_delta_unique_genes <- sankeydf_enricher_delta %>%
  distinct(Genes) %>%
  mutate(ID = Genes, 
         x = 1, 
         y = row_number())

# Pathways unique
df_delta_unique_pathways <- sankeydf_enricher_delta %>%
  distinct(Pathways) %>%
  mutate(ID = Pathways, 
         x = 2, 
         y = row_number())

# Concatenate df's
nodes_delta <- bind_rows(df_delta_unique_genes, df_delta_unique_pathways) %>% 
  select(ID) %>% as.data.frame() %>% 
  mutate(ID_nr = row_number() -1) 
nodes_delta
```

Edges: LBW
```{r}
edges_delta <- sankeydf_enricher_delta %>%
  mutate(gene_type = Genes) %>% 
  left_join(nodes_delta, by = c("Genes" = "ID")) %>% 
  left_join(nodes_delta, by = c("Pathways" = "ID")) %>% 
  mutate(source = ID_nr.x,
         target = ID_nr.y,
         value = -log10(`Adjusted P-value`)) %>% 
  select(source, target, value, gene_type)

edges_delta
```

Create sankey plot: LBW
```{r}
# Create the network graph
network_delta <- sankeyNetwork(
  Links = edges_delta,
  Nodes = nodes_delta,
  Source = "source",
  Target = "target",
  Value = "value",
  NodeID = "ID",
  LinkGroup = "gene_type", # hash out to color nodes
  fontSize = 70,
  nodeWidth = 40,
  NodeGroup = NULL, # hash out to color nodes
  sinksRight = FALSE,
  colourScale = JS("d3.scaleOrdinal(d3.schemeCategory20);"),
  iterations = 0
)

network_delta

# Save file paths
savesankey_html_delta <- paste(path, "pathway_analysis/sankeyplot/04_sankeyplot_topgenes_pathways_A5_deltachanges_proteincoding_SAT_rnaseq.html", sep = "/")
savesankey_jpeg_delta <- paste(path, "pathway_analysis/sankeyplot/04_sankeyplot_topgenes_pathways_A5_deltachanges_proteincoding_SAT_rnaseq.jpeg", sep = "/")

# Save network as html
saveNetwork(network_delta, savesankey_html_delta)

# Save network as png
webshot(savesankey_html_delta, savesankey_jpeg_delta, zoom = 3,
        vwidth = 3000,
        vheight = 2000)
```

# A5: Delta 2
## Sankey plot
```{r}
# Make Sankey df
sankeydf_enricher_delta2 <- enricher_delta %>%
  filter(`Adjusted P-value` < 0.05,
         Term == c("Metabolism Of Steroids R-HSA-8957322", "Muscle Contraction R-HSA-397014")) %>% 
  select(Term, Genes,`Adjusted P-value`) %>% 
  separate_rows(Genes, sep = ";") %>% 
  separate(Term, into = c("Pathways"), sep = "R-HSA") 
sankeydf_enricher_delta2
```


Nodes: LBW
```{r}
# Genes unique
df_delta_unique_genes2 <- sankeydf_enricher_delta2 %>%
  distinct(Genes) %>%
  mutate(ID = Genes, 
         x = 1, 
         y = row_number())

# Pathways unique
df_delta_unique_pathways2 <- sankeydf_enricher_delta2 %>%
  distinct(Pathways) %>%
  mutate(ID = Pathways, 
         x = 2, 
         y = row_number())

# Concatenate df's
nodes_delta2 <- bind_rows(df_delta_unique_genes2, df_delta_unique_pathways2) %>% 
  select(ID) %>% as.data.frame() %>% 
  mutate(ID_nr = row_number() -1) 
nodes_delta2
```

Edges: LBW
```{r}
edges_delta2 <- sankeydf_enricher_delta2 %>%
  mutate(gene_type = Genes) %>% 
  left_join(nodes_delta2, by = c("Genes" = "ID")) %>% 
  left_join(nodes_delta2, by = c("Pathways" = "ID")) %>% 
  mutate(source = ID_nr.x,
         target = ID_nr.y,
         value = -log10(`Adjusted P-value`)) %>% 
  select(source, target, value, gene_type)

edges_delta2
```

Create sankey plot: LBW
```{r}
# Create the network graph
network_delta2 <- sankeyNetwork(
  Links = edges_delta2,
  Nodes = nodes_delta2,
  Source = "source",
  Target = "target",
  Value = "value",
  NodeID = "ID",
  LinkGroup = "gene_type", # hash out to color nodes
  fontSize = 70,
  nodeWidth = 40,
  NodeGroup = NULL, # hash out to color nodes
  sinksRight = FALSE,
  colourScale = JS("d3.scaleOrdinal(d3.schemeCategory20);"),
  iterations = 0
)

network_delta2

# Save file paths
savesankey_html_delta2 <- paste(path, "pathway_analysis/sankeyplot/04_sankeyplot_topgenes_pathways_A5_deltachanges_proteincoding_SAT_rnaseq_2.html", sep = "/")
savesankey_jpeg_delta2 <- paste(path, "pathway_analysis/sankeyplot/04_sankeyplot_topgenes_pathways_A5_deltachanges_proteincoding_SAT_rnaseq_2.jpeg", sep = "/")

# Save network as html
saveNetwork(network_delta2, savesankey_html_delta2)

# Save network as png
webshot(savesankey_html_delta2, savesankey_jpeg_delta2, zoom = 3,
        vwidth = 3000,
        vheight = 2000)
```


# A10: LBWwoNAFLD
## Sankey plot
```{r}
# Make Sankey df
sankeydf_enricher_lbwwoNAFLD <- enricher_lbwwoNAFLD %>%
  filter(`Adjusted P-value` < 0.05) %>% 
  select(Term, Genes,`Adjusted P-value`) %>% 
  separate_rows(Genes, sep = ";") %>% 
  separate(Term, into = c("Pathways"), sep = "R-HSA") 
sankeydf_enricher_lbwwoNAFLD
```


Nodes: LBWwoNAFLD
```{r}
# Genes unique
df_lbwwoNAFLD_unique_genes <- sankeydf_enricher_lbwwoNAFLD %>%
  distinct(Genes) %>%
  mutate(ID = Genes, 
         x = 1, 
         y = row_number())

# Pathways unique
df_lbwwoNAFLD_unique_pathways <- sankeydf_enricher_lbwwoNAFLD %>%
  distinct(Pathways) %>%
  mutate(ID = Pathways, 
         x = 2, 
         y = row_number())

# Concatenate df's
nodes_lbwwoNAFLD <- bind_rows(df_lbwwoNAFLD_unique_genes, df_lbwwoNAFLD_unique_pathways) %>% 
  select(ID) %>% as.data.frame() %>% 
  mutate(ID_nr = row_number() -1) 
nodes_lbwwoNAFLD
```

Edges: LBWwoNAFLD
```{r}
edges_lbwwoNAFLD <- sankeydf_enricher_lbwwoNAFLD %>%
  mutate(gene_type = Genes) %>% 
  left_join(nodes_lbwwoNAFLD, by = c("Genes" = "ID")) %>% 
  left_join(nodes_lbwwoNAFLD, by = c("Pathways" = "ID")) %>% 
  mutate(source = ID_nr.x,
         target = ID_nr.y,
         value = -log10(`Adjusted P-value`)) %>% 
  select(source, target, value, gene_type)

edges_lbwwoNAFLD
```

Create sankey plot: LBWwoNAFLD
```{r}
# Create the network graph
network_lbwwoNAFLD <- sankeyNetwork(
  Links = edges_lbwwoNAFLD,
  Nodes = nodes_lbwwoNAFLD,
  Source = "source",
  Target = "target",
  Value = "value",
  NodeID = "ID",
  LinkGroup = "gene_type", # hash out to color nodes
  fontSize = 70,
  nodeWidth = 40,
  NodeGroup = NULL, # hash out to color nodes
  sinksRight = FALSE,
  colourScale = JS("d3.scaleOrdinal(d3.schemeCategory20);"),
  iterations = 0
)

network_lbwwoNAFLD

# Save file paths
savesankey_html_lbwwoNAFLD <- paste(path, "pathway_analysis/sankeyplot/04_sankeyplot_topgenes_pathways_A10_LBWwoNAFLD_BvsA_proteincoding_SAT_rnaseq.html", sep = "/")
savesankey_jpeg_lbwwoNAFLD <- paste(path, "pathway_analysis/sankeyplot/04_sankeyplot_topgenes_pathways_A10_LBWwoNAFLD_BvsA_proteincoding_SAT_rnaseq.jpeg", sep = "/")

# Save network as html
saveNetwork(network_lbwwoNAFLD, savesankey_html_lbwwoNAFLD)

# Save network as png
webshot(savesankey_html_lbwwoNAFLD, savesankey_jpeg_lbwwoNAFLD, zoom = 3,
        vwidth = 3000,
        vheight = 2000)
```


# A11: LBWwNAFLD
## Sankey plot
```{r}
# Make Sankey df
sankeydf_enricher_lbwwNAFLD <- enricher_lbwwNAFLD %>%
  filter(`Adjusted P-value` < 0.05) %>% 
  select(Term, Genes,`Adjusted P-value`) %>% 
  separate_rows(Genes, sep = ";") %>% 
  separate(Term, into = c("Pathways"), sep = "R-HSA") 
sankeydf_enricher_lbwwNAFLD
```


Nodes: LBWwNAFLD
```{r}
# Genes unique
df_lbwwNAFLD_unique_genes <- sankeydf_enricher_lbwwNAFLD %>%
  distinct(Genes) %>%
  mutate(ID = Genes, 
         x = 1, 
         y = row_number())

# Pathways unique
df_lbwwNAFLD_unique_pathways <- sankeydf_enricher_lbwwNAFLD %>%
  distinct(Pathways) %>%
  mutate(ID = Pathways, 
         x = 2, 
         y = row_number())

# Concatenate df's
nodes_lbwwNAFLD <- bind_rows(df_lbwwNAFLD_unique_genes, df_lbwwNAFLD_unique_pathways) %>% 
  select(ID) %>% as.data.frame() %>% 
  mutate(ID_nr = row_number() -1) 
nodes_lbwwNAFLD
```

Edges: LBWwNAFLD
```{r}
edges_lbwwNAFLD <- sankeydf_enricher_lbwwNAFLD %>%
  mutate(gene_type = Genes) %>% 
  left_join(nodes_lbwwNAFLD, by = c("Genes" = "ID")) %>% 
  left_join(nodes_lbwwNAFLD, by = c("Pathways" = "ID")) %>% 
  mutate(source = ID_nr.x,
         target = ID_nr.y,
         value = -log10(`Adjusted P-value`)) %>% 
  select(source, target, value, gene_type)

edges_lbwwNAFLD
```

Create sankey plot: LBWwNAFLD
```{r}
# Create the network graph
network_lbwwNAFLD <- sankeyNetwork(
  Links = edges_lbwwNAFLD,
  Nodes = nodes_lbwwNAFLD,
  Source = "source",
  Target = "target",
  Value = "value",
  NodeID = "ID",
  LinkGroup = "gene_type", # hash out to color nodes
  fontSize = 70,
  nodeWidth = 40,
  NodeGroup = NULL, # hash out to color nodes
  sinksRight = FALSE,
  colourScale = JS("d3.scaleOrdinal(d3.schemeCategory20);"),
  iterations = 0
)

network_lbwwNAFLD

# Save file paths
savesankey_html_lbwwNAFLD <- paste(path, "pathway_analysis/sankeyplot/04_sankeyplot_topgenes_pathways_A11_LBWwNAFLD_BvsA_proteincoding_SAT_rnaseq.html", sep = "/")
savesankey_jpeg_lbwwNAFLD <- paste(path, "pathway_analysis/sankeyplot/04_sankeyplot_topgenes_pathways_A11_LBWwNAFLD_BvsA_proteincoding_SAT_rnaseq.jpeg", sep = "/")

# Save network as html
saveNetwork(network_lbwwNAFLD, savesankey_html_lbwwNAFLD)

# Save network as png
webshot(savesankey_html_lbwwNAFLD, savesankey_jpeg_lbwwNAFLD, zoom = 3,
        vwidth = 3000,
        vheight = 2000)
```


# A12: Delta: LBW w/o NAFLD vs NBW
## Sankey plot
```{r}
# Make Sankey df
sankeydf_enricher_deltaLBWwoNAFLDvsNBW <- enricher_deltaLBWwoNAFLDvsNBW %>%
  filter(`Adjusted P-value` < 0.05) %>% 
  select(Term, Genes,`Adjusted P-value`) %>% 
  separate_rows(Genes, sep = ";") %>% 
  separate(Term, into = c("Pathways"), sep = "R-HSA") 
sankeydf_enricher_deltaLBWwoNAFLDvsNBW
```


Nodes: LBW
```{r}
# Genes unique
df_deltaLBWwoNAFLDvsNBW_unique_genes <- sankeydf_enricher_deltaLBWwoNAFLDvsNBW %>%
  distinct(Genes) %>%
  mutate(ID = Genes, 
         x = 1, 
         y = row_number())

# Pathways unique
df_deltaLBWwoNAFLDvsNBW_unique_pathways <- sankeydf_enricher_deltaLBWwoNAFLDvsNBW %>%
  distinct(Pathways) %>%
  mutate(ID = Pathways, 
         x = 2, 
         y = row_number())

# Concatenate df's
nodes_deltaLBWwoNAFLDvsNBW <- bind_rows(df_deltaLBWwoNAFLDvsNBW_unique_genes, df_deltaLBWwoNAFLDvsNBW_unique_pathways) %>% 
  select(ID) %>% as.data.frame() %>% 
  mutate(ID_nr = row_number() -1) 
nodes_deltaLBWwoNAFLDvsNBW
```

Edges: LBW
```{r}
edges_deltaLBWwoNAFLDvsNBW <- sankeydf_enricher_deltaLBWwoNAFLDvsNBW %>%
  mutate(gene_type = Genes) %>% 
  left_join(nodes_deltaLBWwoNAFLDvsNBW, by = c("Genes" = "ID")) %>% 
  left_join(nodes_deltaLBWwoNAFLDvsNBW, by = c("Pathways" = "ID")) %>% 
  mutate(source = ID_nr.x,
         target = ID_nr.y,
         value = -log10(`Adjusted P-value`)) %>% 
  select(source, target, value, gene_type)

edges_deltaLBWwoNAFLDvsNBW
```

Create sankey plot: LBW
```{r}
# Create the network graph
network_deltaLBWwoNAFLDvsNBW <- sankeyNetwork(
  Links = edges_deltaLBWwoNAFLDvsNBW,
  Nodes = nodes_deltaLBWwoNAFLDvsNBW,
  Source = "source",
  Target = "target",
  Value = "value",
  NodeID = "ID",
  LinkGroup = "gene_type", # hash out to color nodes
  fontSize = 70,
  nodeWidth = 40,
  NodeGroup = NULL, # hash out to color nodes
  sinksRight = FALSE,
  colourScale = JS("d3.scaleOrdinal(d3.schemeCategory20);"),
  iterations = 0
)

network_deltaLBWwoNAFLDvsNBW

# Save file paths
savesankey_html_deltaLBWwoNAFLDvsNBW <- paste(path, "pathway_analysis/sankeyplot/04_sankeyplot_topgenes_pathways_A12_deltachanges_proteincoding_SAT_rnaseq.html", sep = "/")
savesankey_jpeg_deltaLBWwoNAFLDvsNBW <- paste(path, "pathway_analysis/sankeyplot/04_sankeyplot_topgenes_pathways_A12_deltachanges_proteincoding_SAT_rnaseq.jpeg", sep = "/")

# Save network as html
saveNetwork(network_deltaLBWwoNAFLDvsNBW, savesankey_html_deltaLBWwoNAFLDvsNBW)

# Save network as png
webshot(savesankey_html_deltaLBWwoNAFLDvsNBW, savesankey_jpeg_deltaLBWwoNAFLDvsNBW, zoom = 3,
        vwidth = 3000,
        vheight = 2000) 
```


# A13: Delta: LBW w/ NAFLD vs NBW
## Sankey plot
```{r}
# Make Sankey df
sankeydf_enricher_deltaLBWwNAFLDvsNBW <- enricher_deltaLBWwNAFLDvsNBW %>%
  filter(`Adjusted P-value` < 0.05) %>% 
  select(Genes, Term,`Adjusted P-value`) %>% 
  separate_rows(Genes, sep = ";") %>% 
  separate(Term, into = c("Pathways"), sep = "R-HSA") 
sankeydf_enricher_deltaLBWwNAFLDvsNBW
```

No significant overlaps after FDR adjustment


