---
title: "Multi-Omics Integration Analysis"
subtitle: "Multi-Omics Data Integration Analyses w/ MOFA2"
author: "Sofie Olund Villumsen"
date: "2023"
output: 
  html_document:
    toc: true
    toc_depth: 2
---

# Initial settings
Clear workspace
```{r echo=FALSE}
rm(list = ls()) 
```

Load packages
```{r message=FALSE, warning=FALSE}
library(tidyverse)
library(MOFA2)
library(compositions)
```

Set wd
```{r setup}
path <- "L:/LovbeskyttetMapper/HCOF Stem Cells/Sofie_Databehandling/github/COFpaper_omics_adiposetissue/Multi-omics"

knitr::opts_knit$set(root.dir = path)
```

Load data
```{r message=FALSE}
rnaseq_sat_data_visitAB  <- read_csv(paste(path, "data/01_data_genecounts_proteincoding_rnaseq_sat_reducedsamples_visitAB.csv", sep="/"))

metabolomics_data_visitAB  <- read_csv(paste(path, "data/01_data_abundances_metabolomics_allsamples_visitAB.csv", sep="/")) 

lipidomics_data_visitAB  <- read_csv(paste(path, "data/01_data_abundances_lipidomics_allsamples_visitAB.csv", sep="/")) 
```

Create matrices
```{r}
# RNAseq
rnaseq_sat_data_visitAB_ <- rnaseq_sat_data_visitAB[2:ncol(rnaseq_sat_data_visitAB)] %>% as.matrix() 
rownames(rnaseq_sat_data_visitAB_) <- rnaseq_sat_data_visitAB$gene_id

# Metabolomics
metabolomics_data_visitAB_ <- metabolomics_data_visitAB %>% 
  pivot_longer(cols = -sample_id, names_to = "Metabolites",values_to = "Value") %>% 
  pivot_wider(names_from = sample_id, values_from = Value) %>% 
  select(everything(), -Metabolites) %>% 
  as.matrix()
rownames(metabolomics_data_visitAB_) <- colnames(metabolomics_data_visitAB[2:ncol(metabolomics_data_visitAB)])

# Lipidomics
lipidomics_data_visitAB_ <- lipidomics_data_visitAB %>% 
  pivot_longer(cols = -sample_id, names_to = "Lipids",values_to = "Value") %>% 
  pivot_wider(names_from = sample_id, values_from = Value) %>% 
  select(everything(), -Lipids) %>% 
  as.matrix()
rownames(lipidomics_data_visitAB_) <- colnames(lipidomics_data_visitAB[2:ncol(lipidomics_data_visitAB)])
```


# Define data options for feature selection
```{r}
# Feature selection of genes (RNA-seq) based on variance. Features with a variance below the threshold of 0.4 are removed. Calculating the variance of each gene (row) and creating a logical vector indicating genes having a variance greater than 0.4.
rnaseq_sat_data_visitAB_filtered_data <- rnaseq_sat_data_visitAB_[apply(rnaseq_sat_data_visitAB_, 1, var) > 0.4, ]
dim(rnaseq_sat_data_visitAB_filtered_data)
```

# Check for normalization
## RNA seq
```{r}
# Histogram of total distribution
data_rnaseq <-as.matrix(rnaseq_sat_data_visitAB_filtered_data)
hist(data_rnaseq , prob=T)
curve(dnorm(x,mean(data_rnaseq),sd(data_rnaseq)),add=T,col="red")
```

### Log transformation and scaling
```{r}
# RNA-seq data has skewed distribution. Apply centered log transformation (transpose matrix, as the input format is rows=samples, columns=variables)
rnaseq_sat_data_visitAB_filtered_data_clr <- t( clr( t(rnaseq_sat_data_visitAB_filtered_data) ) )

# Histogram of total distribution
rna_matrix <-as.matrix(rnaseq_sat_data_visitAB_filtered_data_clr)
hist(rna_matrix , prob=T)
curve(dnorm(x,mean(rna_matrix),sd(rna_matrix)),add=T,col="red")

# Check mean of each gene (should equal zero)
##rowMeans(rna_matrix)

# Check SD of each gene (should equal one)
##apply(rna_matrix,1,sd)

# Scale data as the SD of the normalized data is != 1
rnaseq_sat_data_visitAB_filtered_data_clr_scaled <- t( scale( t( rnaseq_sat_data_visitAB_filtered_data_clr ) ) )

# Histogram of total distribution
rna_norm_scaled_matrix <-as.matrix(rnaseq_sat_data_visitAB_filtered_data_clr_scaled)
hist(rna_norm_scaled_matrix , prob=T)
curve(dnorm(x,mean(rna_norm_scaled_matrix),sd(rna_norm_scaled_matrix)),add=T,col="red")

# Histogram of individual distribution of each gene (row)
##rownum = x
##hist(rna_norm_scaled_matrix[,rownum] , prob=T)
##curve(dnorm(x,mean(rna_norm_scaled_matrix[,rownum]),sd(rna_norm_scaled_matrix[,rownum])),add=T,col="red")
```



## Metabolomics
```{r}
# Histogram of total distribution
metabolomics_matrix <- as.matrix(metabolomics_data_visitAB_)
hist(metabolomics_matrix , prob=T)
curve(dnorm(x,mean(metabolomics_matrix),sd(metabolomics_matrix)),add=T,col="red")

# Check mean of each gene (should equal zero)
##rowMeans(metabolomics_matrix)

# Check SD of each gene (should equal one)
##apply(metabolomics_matrix,1,sd)
```



## Lipidomics
```{r}
# Histogram of total distribution
lipidomics_matrix <-as.matrix(lipidomics_data_visitAB_)
hist(lipidomics_matrix , prob=T)
curve(dnorm(x,mean(lipidomics_matrix),sd(lipidomics_matrix)),add=T,col="red")

# Check mean of each gene (should equal zero)
##rowMeans(lipidomics_matrix)

# Check SD of each gene (should equal one)
##apply(lipidomics_matrix,1,sd)
```


# Add missing sample columns to RNA-seq data
```{r}
# RNA-seq data
## Identify missing columns
missing_cols <- setdiff(metabolomics_data_visitAB$sample_id, names(rnaseq_sat_data_visitAB[2:ncol(rnaseq_sat_data_visitAB)]))

## Get the column order of metabolomics_data_visitAB
sample_order <- metabolomics_data_visitAB$sample_id

## Add missing columns 
rnaseq_sat_data_visitAB_allsamples <- as_tibble(rnaseq_sat_data_visitAB_filtered_data) %>% add_column(!!!set_names(rep(list(NA), length(missing_cols)), missing_cols), .before = 1) %>% 
  select(all_of(sample_order)) %>% as.matrix()
rownames(rnaseq_sat_data_visitAB_allsamples) <- rownames(rnaseq_sat_data_visitAB_filtered_data)
```



# 3-layer multi-omics integration: Create the MOFA object
```{r}
# Create a list containing the 3 omics matrices
omics3layer_matrix_list <- list(RNAseq = rnaseq_sat_data_visitAB_allsamples, 
                                Lipidomics = lipidomics_data_visitAB_,
                                Metabolomics = metabolomics_data_visitAB_)
lapply(omics3layer_matrix_list, dim)
```
```{r}
# Create MOFA object
MOFAobject_3layer <- create_mofa(omics3layer_matrix_list)
```

## Plot the data overview
```{r}
plot_data_overview(MOFAobject_3layer, colors = c("#336699", "#CC6600", "#F5BD1E")) 
```



## Loop defining options for 10 models with factors 1-10
```{r}
# Define data options
data_opts <- get_default_data_options(MOFAobject_3layer)

# Define training options
train_opts <- get_default_training_options(MOFAobject_3layer)

# Loop over num_factors
for (num_factors in 1:10) {
  # Define model options
  model_opts <- get_default_model_options(MOFAobject_3layer)
  model_opts$num_factors <- num_factors

  # Build and train the MOFA object
  MOFAobject <- prepare_mofa(
    object = MOFAobject_3layer,
    data_options = data_opts,
    model_options = model_opts,
    training_options = train_opts
  )

  # Train the MOFA model
  file_name <- paste0("data/02_MOFAmodel_3layer_", num_factors, "factor.hdf5")
  outfile <- file.path(path, file_name)
  MOFAobject.trained <- run_mofa(MOFAobject, outfile, use_basilisk = TRUE)
}
```



