---
title: "Multi-Omics Integration Analysis"
subtitle: "Downstream multi-Omics Data Integration Analyses w/ MOFA2"
author: "Sofie Olund Villumsen"
date: "2023"
output: 
  html_document:
    toc: true
    toc_depth: 2
---

# Initial settings
Clear workspace
```{r echo=FALSE}
rm(list = ls()) 
```

Load packages
```{r message=FALSE, warning=FALSE}
library(tidyverse)
library(MOFA2)
library(correlation) 
library(ggside)
```

Set wd
```{r setup}
path <- "L:/LovbeskyttetMapper/HCOF Stem Cells/Sofie_Databehandling/github/COFpaper_omics_adiposetissue/Multi-omics"

knitr::opts_knit$set(root.dir = path)
```

Load data
```{r message=FALSE}
# Clinical data/meta data
clinicaldata_visitAB <- read_csv(paste(path, "data/01_data_clinicaldata_allsamples_visitAB.csv", sep="/")) %>% 
  select(everything(), -id) 
```

# Load 3 layer omics models
```{r}
model_1factor <- load_model(paste(path, "data/02_MOFAmodel_3layer_1factor.hdf5", sep="/"))
model_2factor <- load_model(paste(path, "data/02_MOFAmodel_3layer_2factor.hdf5", sep="/"))
model_3factor <- load_model(paste(path, "data/02_MOFAmodel_3layer_3factor.hdf5", sep="/"))
model_4factor <- load_model(paste(path, "data/02_MOFAmodel_3layer_4factor.hdf5", sep="/"))
model_5factor <- load_model(paste(path, "data/02_MOFAmodel_3layer_5factor.hdf5", sep="/"))
model_6factor <- load_model(paste(path, "data/02_MOFAmodel_3layer_6factor.hdf5", sep="/"))
model_7factor <- load_model(paste(path, "data/02_MOFAmodel_3layer_7factor.hdf5", sep="/"))
model_8factor <- load_model(paste(path, "data/02_MOFAmodel_3layer_8factor.hdf5", sep="/"))
model_9factor <- load_model(paste(path, "data/02_MOFAmodel_3layer_9factor.hdf5", sep="/"))
model_10factor <- load_model(paste(path, "data/02_MOFAmodel_3layer_10factor.hdf5", sep="/"))
```


# Overview of data
```{r}
# Overview plot, which shows how many views (rows) and how many groups (columns) exist, what are their corresponding dimensions and how many missing information they have (grey bars).
dataoverview_obj <- plot_data_overview(model_1factor, 
                                       colors = c("#63A0CB", "#FFA657", "#6CBD6C")) 

# Save the object as a PDF file
tiff(file = paste(path, "results/03_dataoverview_3layers_visitAB_multiomics.tiff", sep="/"),
    width = 10,
    height = 6,
    units = "cm",
    res = 500)

print(dataoverview_obj)
dev.off()

dataoverview_obj
```


# Variance decomposition
```{r}
# The first step in the MOFA analysis is to quantify the amount of variance explained (R2) by each factor in each data modality.

f01 <- model_1factor@cache$variance_explained$r2_total[[1]] # group 1
f02 <- model_2factor@cache$variance_explained$r2_total[[1]] # group 1
f03 <- model_3factor@cache$variance_explained$r2_total[[1]] # group 1
f04 <- model_4factor@cache$variance_explained$r2_total[[1]] # group 1
f05 <- model_5factor@cache$variance_explained$r2_total[[1]] # group 1
f06 <- model_6factor@cache$variance_explained$r2_total[[1]] # group 1
f07 <- model_7factor@cache$variance_explained$r2_total[[1]] # group 1
f08 <- model_8factor@cache$variance_explained$r2_total[[1]] # group 1
f09 <- model_9factor@cache$variance_explained$r2_total[[1]] # group 1
f10 <- model_10factor@cache$variance_explained$r2_total[[1]] # group 1

data_factors <- data.frame(f01,f02,f03,f04,f05,f06,f07,f08,f09,f10) %>% 
  mutate(omics = rownames(.)) %>% 
  pivot_longer(cols = starts_with("f"), names_to = "factors" , values_to = "expl_var") 
data_factors
```

## Plot proportion of variance explained by omics view
```{r}
explainedvar_plot <- ggplot(data = data_factors, aes(x = factors, y = expl_var, col = omics, group = omics)) +
  geom_point(size = 2) +
  geom_line(size = 0.8) +
  # Customize the plot aesthetics
  labs(x = "Factors", y = "Explained variance (%)", title = "Proportion of variance explained by omics view") +
  theme_minimal() +
  scale_color_manual(values = c("#63A0CB", "#6CBD6C", "#FFA657"),
                     breaks = c("RNAseq", "Metabolomics", "Lipidomics"))


# Save the object as a PDF file
png(file = paste(path, "results/03_explainedvarianceplot_3layers_visitAB_multiomics.png", sep="/"),
    width = 13,
    height = 8,
    units = "cm",
    res = 300)

print(explainedvar_plot)
dev.off()

explainedvar_plot
```

## Variance explained for every factor (f1-10) in per view and group
```{r}
factor10_varmatrix <- model_10factor@cache$variance_explained$r2_per_factor[[1]] %>% # group 1  
  as.data.frame()
factor10_varmatrix

factor10_df <- as_tibble(factor10_varmatrix) %>% 
  mutate(Factors = rownames(factor10_varmatrix))
factor10_df
```

## Variance explained for every factor (f1-2) in per view and group
```{r}
# Two factors are chosen, with a cutoff of minimum 2% in each omics view for each factor (biological meaning)
factor2_varmatrix <- model_2factor@cache$variance_explained$r2_per_factor[[1]] %>% # group 1  
  as.data.frame() 

factor2_df <- as_tibble(factor2_varmatrix) %>% 
  mutate(Factors = rownames(factor2_varmatrix))
factor2_df
```

## Total variande explained for each factor
```{r}
total_factorvariance <- model_2factor@cache$variance_explained$r2_per_factor[[1]] %>% # group 1  
  as.data.frame() %>% 
  mutate(`Total variance` = rowSums(across(RNAseq:Lipidomics)),
         Factor = rownames(.)) %>% 
  select(Factor, `Total variance`)
total_factorvariance
```


## Add metadata to the model (f1-2)
```{r}
clinicaldata_visitAB_ <- clinicaldata_visitAB  %>% 
  rename(sample = sample_id) %>% 
  mutate(bwcat_visit = paste(.$bwcat, .$visit, sep="_"))

samples_metadata(model_2factor) <- clinicaldata_visitAB_
head(model_2factor@samples_metadata, n=25)
```

# Visualizations

## Plot variance explained pr factor
```{r}
# Reshape the data frame to long format
df_long <- pivot_longer(factor2_df, cols = -Factors, names_to = "Omics_Type", values_to = "Value")

# Specify the desired order for the Omics_Type variable
df_long$Omics_Type <- factor(df_long$Omics_Type, levels = c("Metabolomics", "Lipidomics", "RNAseq", "Total variance"))

df_long$Factors <- factor(df_long$Factors, levels = c("Factor1", "Factor2"))

# Create the ggplot
var_expl_prfactor <- ggplot(df_long, aes(x = Factors, y = Value, fill = Omics_Type)) +
  geom_bar(stat = "identity", position = "dodge", width = 0.7) +
  labs(x = "Factor \n(multi-omics feature combination)", y = "Variance explained (%)", fill = "Omics Type") +
  scale_fill_manual(values = c("RNAseq" = "#63A0CB", "Metabolomics" = "#6CBD6C", "Lipidomics" = "#FFA657", "Total variance" = "lightgrey")) +
  theme_classic() +
  theme(
    # Adjust the size of the legend title
    legend.title = element_text(size = 20),
    # Adjust the size of legend text
    legend.text = element_text(size = 17),
    # Adjust the size of axis text
    axis.text.x = element_text(size = 17),
    axis.text.y = element_text(size = 17),
    axis.title.x = element_text(size = 20),
    axis.title.y = element_text(size = 20)
  )

var_expl_prfactor

# Save the object as a PDF file
tiff(file = paste(path, "results/03_explainedvarianceprfactorplot_3layers_visitAB_multiomics.tiff", sep="/"),
    width = 25,
    height = 15,
    units = "cm",
    res = 500)

print(var_expl_prfactor)
dev.off()


```


## Plot variance estimates
```{r}
plot_variance_explained(model_2factor, x="view", y="factor")
```


## Plot variance explained pr. omics view
```{r}
plot_variance_explained(model_2factor, x="group", y="factor", plot_total = T)[[2]] 
```

## Plot single factors
```{r}
model2factor_df <- as.data.frame(model_2factor@expectations$Z$group1) %>% 
  mutate(sample = rownames(model_2factor@expectations$Z$group1)) %>% 
  left_join(clinicaldata_visitAB_, by = join_by(sample))
```

### Bwcat and visit
```{r}
plot_factor(model_2factor, 
            factors = 1:2, 
            color_by = "bwcat_visit",
            shape_by = "visit"
)
```
### Birthweight
```{r}
plot_factor(model_2factor, 
  factor = 1:2,
  color_by = "Birth weight - NAFLD status"
)
```
## Boxplot 1: bwcat and visit
```{r}
ggplot(model2factor_df, aes(bwcat_visit, Factor1)) + 
  geom_boxplot(aes(fill = factor(bwcat_visit)), alpha = 0.5) +
  theme_classic() + 
  geom_dotplot(binaxis='y', stackdir='center', position=position_dodge(1), binwidth =0.1) 
#+ ylim(-5.5, 5)


ggplot(model2factor_df, aes(bwcat_visit, Factor2))  + 
  geom_boxplot(aes(fill = factor(bwcat_visit)), alpha = 0.5) +
  theme_classic() + 
  geom_dotplot(binaxis='y', stackdir='center', position=position_dodge(1), binwidth =0.05) 
#+  ylim(-5.5, 5)

```


## Boxplot 2: bwcat and visit
```{r}
p <- plot_factor(model_2factor, 
  factors = c(1,2),
  group_by = "bwcat_visit",
  color_by = "bwcat_visit",
  dot_size = 3,        # change dot size
  dodge = T,           # dodge points with different colors
  legend = T,          # remove legend
  add_violin = F,      # add violin plots,
  violin_alpha = 0.25,  # transparency of violin plots
  add_boxplot = T
) 
  
p
```

## Plot combinations of factors
###  Scatter plot 1: bwcat and visit
```{r}
plot_factors(model_2factor, 
  factors = 1:2,
  color_by = "bwcat_visit",
  dot_size = 4,
  alpha = 0.7)   +
  stat_ellipse(type = "t", level = 0.95) 
```

###  Scatter plot 2: bwcat and visit
```{r}
ggplot(model2factor_df, aes(x = Factor1, y = Factor2, color = bwcat_visit)) +
  geom_point(size = 3, alpha = 0.6) +
  stat_ellipse(size = 0.8, alpha = 0.5) +
  xlab(paste(total_factorvariance$Factor[1], " (" , round(total_factorvariance$`Total variance`[1]), "% combined variation)", sep = "")) + 
  ylab(paste(total_factorvariance$Factor[2], " (" , round(total_factorvariance$`Total variance`[2]), "% combined variation)", sep = "")) + 
  theme_classic() 
```

### Scatter boxplot
```{r}

model2factor_df_scatterboxplot <- model2factor_df %>% 
  mutate(Group = case_when(
    bwcat_visit == "NBW_A" ~ "NBW baseline",
    bwcat_visit == "NBW_B" ~ "NBW COF",
    bwcat_visit == "LBW_A" ~ "LBW baseline",
    bwcat_visit == "LBW_B" ~ "LBW COF"
  )) 

model2factor_df_scatterboxplot$Group <- factor(model2factor_df_scatterboxplot$Group,
                                               levels = c("NBW baseline", "NBW COF", "LBW baseline", "LBW COF"))

# Define custom colors for the groups
custom_colors <- c("NBW baseline" = "#00AFBB", "NBW COF" = "#045F5F","LBW baseline" = "#E7B800", "LBW COF" = "#F87431")

# Scatterboxplot
scatterboxplot <- ggplot(model2factor_df_scatterboxplot, aes(x = Factor1, y = Factor2, fill = Group)) +
  geom_point(aes(color = Group), size = 2) +
  
  stat_ellipse(size = 0.8, aes(color = Group, fill = Group)) +
  xlab(paste("LF1", " (" , round(total_factorvariance$`Total variance`[1]), 
             "% combined variation)", sep = "")) + 
  ylab(paste("LF2", " (" , round(total_factorvariance$`Total variance`[2]), 
             "% combined variation)", sep = "")) + 
  
  geom_xsideboxplot(aes(y =Group), orientation = "y") +
  scale_xsidey_discrete() + 
  geom_ysideboxplot(aes(x = Group), orientation = "x") +
  scale_ysidex_discrete(guide = guide_axis(angle = 90)) +
  
  # Use custom colors for the groups
  scale_fill_manual(values = custom_colors) +
  scale_color_manual(values = custom_colors) +
  
  theme_classic() +
  theme(legend.position = "bottom") +
  theme(ggside.panel.scale = .3,
        ggside.panel.border = element_rect(NA, "white"),
        ggside.panel.grid = element_line("white"),

            # Adjust the size of the legend title
        legend.title = element_text(size = 17),
    # Adjust the size of legend text
    legend.text = element_text(size = 14),
    # Adjust the size of axis text
    axis.title.x = element_text(size = 17),
    axis.title.y = element_text(size = 17),
        
        # Remove group names on boxplots
        axis.text.y = element_blank(),
        axis.text.x = element_blank()
        
        )
scatterboxplot

ggsave(paste(path, "results/03_scatterboxplot_f1f2_visitAB_multiomics.png", sep="/"), plot = scatterboxplot, device = "png", width = 8.5, height = 6)

# TIFF file
tiff(file = paste(path, "results/03_scatterboxplot_f1f2_visitAB_multiomics.tiff", sep="/"), units="in", width=8.5, height=6, res=600, pointsize = 0.0001)
print(scatterboxplot)
dev.off()
```





## Plot feature weights
The weights provide a score for how strong each feature relates to each factor. Features with no association with the factor have values close to zero, while features with strong association with the factor have large absolute values. The sign of the weight indicates the direction of the effect: a positive weight indicates that the feature has higher levels in the cells with positive factor values, and vice versa.


### RNAseq weights
```{r}
plot_weights(model_2factor,
  view = "RNAseq",
  factor = c(1,2),
  nfeatures = 10,     # Number of features to highlight
  scale = T,          # Scale weights from -1 to 1
  abs = F             # Take the absolute value?
)
```

```{r}
rnaseq_weights_plot <- plot_top_weights(model_2factor,
  view = "RNAseq",
  factor = c(1,2),
  nfeatures = 10, 
  abs = T
)

# Save the object as a PDF file
tiff(file = paste(path, "results/03_weightsplot_rnaseq_3layers_visitAB_multiomics.tiff", sep="/"),
    width = 25,
    height = 10,
    units = "cm",
    res = 500,
    pointsize = 0.0001)
print(rnaseq_weights_plot)
dev.off()

rnaseq_weights_plot
```


### Metabolomics weights
```{r}
plot_weights(model_2factor,
  view = "Metabolomics",
  factor = c(1,2),
  nfeatures = 10,     # Number of features to highlight
  scale = T,          # Scale weights from -1 to 1
  abs = F             # Take the absolute value?
)
```

```{r}
metabolomics_weights_plot <- plot_top_weights(model_2factor,
  view = "Metabolomics",
  factor = c(1,2),
  nfeatures = 10
)

# Save the object as a PDF file
tiff(file = paste(path, "results/03_weightsplot_metabolomics_3layers_visitAB_multiomics.tiff", sep="/"),
    width = 30,
    height = 10,
    units = "cm",
    res = 500,
    pointsize = 0.0001)
print(metabolomics_weights_plot)
dev.off()

metabolomics_weights_plot
```



### Lipidomics weights
```{r}
plot_weights(model_2factor,
  view = "Lipidomics",
  factor = c(1,2),
  nfeatures = 10,     # Number of features to highlight
  scale = T,          # Scale weights from -1 to 1
  abs = F             # Take the absolute value?
)
```

```{r}
lipidomics_weights_plot <- plot_top_weights(model_2factor,
  view = "Lipidomics",
  factor = c(1,2),
  nfeatures = 10, 
  abs = F
)

# Save the object as a PDF file
tiff(file = paste(path, "results/03_weightsplot_lipidomics_3layers_visitAB_multiomics.tiff", sep="/"),
    width = 27,
    height = 10,
    units = "cm",
    res = 500,
    pointsize = 0.0001)

print(lipidomics_weights_plot)
dev.off()

lipidomics_weights_plot
```


## Heatmap of patterns in the input data
Observe the coordinated heterogeneity that MOFA captures in the original data, instead of weights. Top features are selected by its weight in the selected factor. By default, samples are ordered according to their corresponding factor value.

### RNAseq heatmap f2
```{r}
heatmap_rnaseq_obj <- plot_data_heatmap(model_2factor,
  view = "RNAseq",         # view of interest
  factor = 2,             # factor of interest
  features = 10,          # number of features to plot (they are selected by weight)
  
  # extra arguments that are passed to the `pheatmap` function
  cluster_rows = TRUE, cluster_cols = TRUE,
  show_rownames = TRUE, show_colnames = TRUE
)


# Save the object as a PDF file
png(file = paste(path, "results/03_heatmap_rnaseq_f2_visitAB_multiomics.png", sep="/"),
    width = 40,
    height = 30,
    units = "cm",
    res = 300)

print(heatmap_rnaseq_obj)
dev.off()

heatmap_rnaseq_obj
```


### Metabolomics heatmap f2
```{r}
heatmap_metabolomics_obj <- plot_data_heatmap(model_2factor,
  view = "Metabolomics",         # view of interest
  factor = 2,             # factor of interest
  features = 10,          # number of features to plot (they are selected by weight)
  
  # extra arguments that are passed to the `pheatmap` function
  cluster_rows = TRUE, cluster_cols = T,
  show_rownames = TRUE, show_colnames = TRUE
)

# Save the object as a PDF file
png(file = paste(path, "results/03_heatmap_metabolomics_f2_visitAB_multiomics.png", sep="/"),
    width = 40,
    height = 30,
    units = "cm",
    res = 300)

print(heatmap_metabolomics_obj)
dev.off()

heatmap_metabolomics_obj
```

### Lipidomics heatmap f2
```{r}
heatmap_lipidomics_obj <- plot_data_heatmap(model_2factor,
  view = "Lipidomics",         # view of interest
  factor = 2,             # factor of interest
  features = 10,          # number of features to plot (they are selected by weight)
  
  # extra arguments that are passed to the `pheatmap` function
  cluster_rows = TRUE, cluster_cols = TRUE,
  show_rownames = TRUE, show_colnames = TRUE
)


# Save the object as a PDF file
png(file = paste(path, "results/03_heatmap_lipidomics_f2_visitAB_multiomics.png", sep="/"),
    width = 40,
    height = 30,
    units = "cm",
    res = 300)

print(heatmap_lipidomics_obj)
dev.off()

heatmap_lipidomics_obj
```

## Linear regression of observations vs factor values
It is useful to add a linear regression estimate to visualize if the relationship between (top) features and factor values is linear.

### RNAseq scatter plots f2
```{r}
plot_data_scatter(model_2factor,
  view = "RNAseq",         # view of interest
  factor = 2, # factor of interest
  features = 10,           # number of features to plot (they are selected by weight)
  add_lm = TRUE,          # add linear regression
  color_by = "bwcat_visit"
)
```

### Metabolomics scatter plots f2
```{r}
plot_data_scatter(model_2factor,
  view = "Metabolomics",         # view of interest
  factor = 2,             # factor of interest
  features = 10,           # number of features to plot (they are selected by weight)
  add_lm = TRUE,          # add linear regression
  color_by = "bwcat_visit"
)
```

### Lipidomics scatter plots f2
```{r}
plot_data_scatter(model_2factor,
  view = "Lipidomics",         # view of interest
  factor = 2,             # factor of interest
  features = 10,           # number of features to plot (they are selected by weight)
  add_lm = TRUE,          # add linear regression
  color_by = "bwcat_visit"
)
```

# Other functionalities

## Correlate with clinical variables
```{r}
clin_var <- model2factor_df %>% 
  mutate(`Birth weight` = `Birth weight - NAFLD status`) %>% 
  select(c("HDL cholesterol", "LDL cholesterol", "Total cholesterol", "F-triglyceride", "GGT", "ALAT", "ASAT", "Hepatic fat", "HOMA-IR", "F-insulin", "F-C-peptide", "F-glucose", "VAT mass", "Total fat mass", "Total lean mass", "BMI", "Weight", "Height", "Birth weight", "NAFLD status", "Visit", "Birth weight category"))

rownames(clin_var) <- model2factor_df$sample

factors <- model2factor_df %>% 
  select(c(Factor1, Factor2))
rownames(factors) <- model2factor_df$sample

# Compute correlation (rho) and p-values (BH) for several pairs of variables at once.
corr_matrix <- correlation(factors, clin_var, method = "pearson", p_adjust = "BH") %>%  as_tibble()

corr_matrix_ <- corr_matrix %>% 
  filter(p<0.01)
corr_matrix_
```

### Eigencorr plot
```{r}
row_order <- colnames(clin_var)

# Create a correlation heatmap (rho)
heatmap_rho <- ggplot(corr_matrix, aes(Parameter1, Parameter2, fill = r)) +
  geom_tile() +
  scale_fill_gradient2(low = "#283A90", mid = "white", high = "#BE2625", midpoint = 0) +
  geom_text(aes(label = ifelse(abs(r) > 0.20, paste(round(r, 2), ifelse(p <= 0.001, "***", ifelse(p <= 0.01, "**", ifelse(p <= 0.05, "*", "")))), "")), color = "black", size=3) +
  labs(title = "Correlation of LFs and clinical variables (Pearsons r)",
       x = "",
       y = "") + 
  theme_classic() +
  scale_y_discrete(limits = row_order)


tiff(file = paste(path, "results/03_eigencorplot_PearsonsR_visitAB_multiomics.tiff", sep="/"), units="in", width=8, height=13, res=300, pointsize = 0.0001)

print(heatmap_rho)

dev.off()

heatmap_rho
```


For further info see: 
https://raw.githack.com/bioFAM/MOFA2_tutorials/master/R_tutorials/downstream_analysis.html 
